### 2023-04-25 manuscript/manuscriptRW.Rmd
update references table 1

---
title: "Tree establishment in models of forest dynamics: a key priority for further research"
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Kristin H. Braziunas,
- Josef	Bruna,
- Samuel Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred J. Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Katarina Merganicova,
- Tobias Mette,
- Xavier Morin,
- Werner Rammer,
- Christopher P.O. Reyer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
    extra_dependencies: "subfig"
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2007, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016, @langan2017, @scheiter2013 
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

```{r, echo=FALSE}
  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries = FALSE)
  library(bibtex)
  options(tinytex.verbose = TRUE)
  #knitr::opts_chunk$set(echo = TRUE, fig.pos = "h", out.extra = "")

```

# Abstract

Tree establishment is a key process in forest dynamics, particularly in
the context of forest resilience and climate change. Models are pivotal
for assessing long term (i.e., decadal to centennial) forest dynamics,
and they have been in use for more than 50 years. However, an assessment
of their ability to accurately represent tree establishment is lacking.
We assess how well current models capture the overall abundance, species
composition and mortality of tree establishment.

Using 15 models and model variants built to capture long-term forest
dynamics at the stand, landscape and global levels, we simulate tree
establishment at 200 sites representing large environmental gradients
across Europe. The results are evaluated against comprehensive data from
unmanaged forests.

Most of the models overestimate ingrowth levels, which is only
compensated by some models with high simulated mortality rates in the
early stages of individual trees dynamics. Simulated species diversity
of ingrowth matches the observed ranges quite well. Models simulating
higher species diversity at the stand level do not feature higher
ingrowth diversity. The effect of light availability on ingrowth levels
is captured better than the effect of temperature and soil moisture, but
patterns are not consistent across models. The ingrowth of *Abies alba*
is captured well by most models across temperature and soil moisture
gradients, whereas ingrowth of *Fagus sylvatica* is generally
underestimated. *Picea abies* ingrowth is generally overestimated,
particularly at colder sites. Increasing complexity in the tree
establishment modules of the models is not related to higher accuracy of
simulated tree establishment. Furthermore, individual model design is
more important than scale (stand, landscape, global) and approach
(empirical, process-based) for accurately capturing tree establishment.

Despite considerable mismatches between simulation results and data in
some instances, it is remarkable that most of the models capture the
essential features of the highly complex process of tree establishment,
while not having been parameterized with such data. We conclude that in
spite of all the uncertainties, much can be gained by evaluating and
refining model establishment processes. This has the potential to render
long-term projections of forest dynamics under changing environmental
conditions much more robust.

\newpage

# Introduction

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbances impacts and changing societal
demands on forests ecosystems services, it is becoming ever more
important to understand how forest structure, composition and function
will change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics were developed over the past decades considering the
impacts of climate on forest dynamics [@bugmann2022]. From these
studies, it is evident that we have a reasonably good understanding how
to model tree growth [@vanclay1997; @bugmann1996], and substantial
efforts have been dedicated to improving the representation of tree
mortality [@bugmann2019; @cailleret2017]. In contrast, tree
establishment is a much less studied process that is often represented
rather coarsely in these models [@walck2011; @leishman1992; @price2001],
which is a clear research gap in the context of climate-induced forest
disturbances and forest resilience.

Tree establishment arises from multiple processes such as pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree establishment is
challenging. Currently, tree establishment processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
of a few simple environmental filters, as done in most forest "gap"
models [@shuman2011] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994].

Overall, models are needed to (1) synthesize existing empirical data and
explore their relationships, (2) assess future tree establishment e.g.
in the context of global change scenarios, and (3) identify the most
important processes that are shaping ecological patterns. Given the
current strategies that are used in models of forest dynamics to
represent tree establishment, model behavior often is prone to problems,
such as very high levels of tree establishment that necessitate excess
mortality at early stages of tree life to simulate correct stand
structure and composition. Also, inadequately high species diversity in
tree establishment may be simulated, which is characteristic of
"classical" forest gap models [@gutiérrez2016], at least as long as the
simulations set-up comprise a multitude of species. Some models use
calibration against local canopy-level data to define appropriate
ingrowth levels, which is likely to strongly hamper the general
applicability of these models e.g. under scenarios of climate change.
Furthermore, correctly capturing the species composition in tree
establishment as a function of the presence of seed trees in the canopy
is often a particular challenge, potentially leading to unrealistic
successional drift in the model, which must be corrected by factors that
are hard to parameterize [e.g., @lischke2006]. A related issue is the
excessive reduction of species diversity due to positive feedback
effects such that eventually just single-species stands remain. This is
sometimes corrected by the incorporation of a low level of seed influx
of all species at all times [@schumacher2006] or by restricting the
number of seeds per species in the seedbank [@lischke2006]. However,
simulated species composition is usually exceedingly sensitive to
assumptions about seed availability, whereas the parameters of such
functions are poorly constrained by field data. Lastly, there is often a
problem with insufficient observational constraints on parameter values
for models that start from very small tree sizes (e.g., 10 cm of height
or even seeds) and track tree development in a process-oriented manner
by considering a multitude of ecological influences, rather than
emphasizing tree establishment into a larger size class [e.g.,
@wehrli2006].

Thus, a focus on the modeling of tree establishment processes is sorely
needed and overdue [@price2001; @walck2011] if we are to make reliable
projections of future forest dynamics, i.e., when the models need to be
operated in extrapolation mode, as well as from a fundamental ecological
point of view for increasing systems understanding. Also in empirical
ecology, tree establishment has not received a lot of attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail [@miina2006]. Yet, little data is available that covers all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Thus, the available data are fragmented,
which constitutes a major problem for model building [@clark1999]. On
the one hand, monitoring on permanent sample plots (such as National
Forest Inventories) often measures tree establishment [@zell2019],
however with a design that captures the rate after the stand initiation
phase [@hallsby2015]. Data from permanent plots, although available
across regions, are highly heterogeneous, rendering their use difficult
in a modeling context [@käber2021]. On the other hand, targeted studies
to measure tree establishment on experimental sites are often limited in
spatial extent, species studied, or the sub-set of processes that are
targeted [@berdanier2016; @collet2006].

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree establishment [@käber2022]. By tree establishment, we refer to
the passing of trees across a specific diameter threshold ("ingrowth").
We included models that are based on a range of "philosophies", from
models operating at the stand to the global scale as well as the range
of models from empirically derived to "process-based" [e.g.,
@bugmann1996; @fabrika2012; @lexer2001; @reyer2014]. Because of the
large variability in tree establishment patterns in nature and the large
number of factors driving this process --- including some that are not
incorporated in most models such as deer browsing --- we do not aim at a
detailed statistical evaluation of each model. Instead, we aim at
evaluating the general ingrowth patterns and magnitudes simulated by the
models, and benchmark the simulated regeneration niche of multiple
species against empirical data along a wide environmental gradient of
temperature, moisture, and light availability [@grubb1977].

More specifically, we aim to answer the following questions: (1) Are
models of forest dynamics capturing accurately tree ingrowth levels,
initial tree species diversity and mortality in the establishment? (2)
Do model traits explain differences in model performance? (3) How well
do the models capture total ingrowth and individual species regeneration
niches across environmental gradients of light availability,
temperature, and soil moisture?

# Material and methods

## Models

Fifteen models of forest dynamics (including two models featuring model
variants) were used to assess simulated tree establishment. They
simulate forest dynamics at the stand, landscape or global scale (Table
\@ref(tab:tableModels)). The approaches used for model construction and
their origin differ strongly, with most of the models featuring a
largely "process-based" approach, whereas two models are based on
formulations derived from the statistical analysis of inventory data
(SIBYLA and xComp). Some of the models largely rely on the original
approach underlying forest gap models (e.g., ForClim 1 and LandClim),
whereas some are based on plant ecophysiological processes (e.g.,
FORMIND and iLand). The different approaches underlying the models have
strongly influenced the formulation of tree establishment processes.

The establishment modules can be categorized as 'regeneration' or
'recruitment' models [@vanclay1994; @konig_tree_2022] (see Table
\@ref(tab:tableModels)). Regeneration models include processes such as
flowering and pollination, seed production, seed dispersal, germination,
and seedling growth, which ultimately lead to the simulated number of
established trees. Recruitment models, in contrast, introduce a number
of new trees with certain characteristics such as biomass or diameter,
without explicitly considering earlier development processes. We can
further distinguish models that feature a feedback in the establishment,
i.e., where the level and species composition of establishment is
influenced by the existing adult tree community via the production of
seeds, seedlings or saplings, from models that do not contain such a
feedback.

The starting point for tree establishment in the models differs as well,
ranging from seedbank, seed or seedling to sapling (i.e., trees much
larger than 10 cm height, often ca. 2-3 m tall). Models that start from
seed need to include a larger number of ecological processes such as
germination and survival of young seedlings, whereas models that start
from saplings have to aggregate via parameterizations several ecological
processes which are not treated explicitly. This later approach reduces
model complexity but comes at the cost of blurring process
representation.

In the models used in this study, the overall complexity in the
establishment modules varies considerably. Following @bugmann2022, we
can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS,
ForClim and TreeMig) as having rather high complexity in their
establishment modules (mean establishment formulation complexity across
all processes \>0, Table \@ref(tab:tableModels)). Two models, ForClim
and ForCEEPS, were used here with two alternative variants of
establishment. ForClim variant 1 [@bugmann1996] is based on a
recruitment module that adheres closely to the concept introduced by
@botkin1972, whereas ForClim variant 11 is adopting a different approach
where individual species properties and their relation to the
environment are important [@huber2020]. These two model variants allow
us to evaluate a more process-based and complex module (variant 11)
against a simple module (variant 1), while the rest of the model
structure is identical. Similarly, the two ForCEEPS variants allow us to
isolate the importance of the canopy feedback (i.e., simulated actual
composition and relative abundance of species in the plot) via seed
trees for the quantity and quality (e.g., diversity and composition) of
simulated ingrowth, as one variant includes this feedback whereas the
other does not.

```{r tableModels, echo=FALSE,  message=FALSE, warning=FALSE}
library(knitr)

modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))
modelDF$Sampling <- NULL

#footnotes
names(modelDF)[2] <- paste0(names(modelDF)[2], "(a)")
names(modelDF)[3] <- paste0(names(modelDF)[3], "(b)")
names(modelDF)[4] <- paste0(names(modelDF)[4], "(c)")
names(modelDF)[5] <- paste0(names(modelDF)[5], "(d)")
names(modelDF)[6] <- paste0(names(modelDF)[6], "(e)")
names(modelDF)[11] <- paste0(names(modelDF)[11],"(f)")

kable(modelDF, 
      format = "latex", booktabs = TRUE,
      caption = "Models included and their characteristics.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |> 
      kableExtra::column_spec(column = 2, width = "3em") |> 
      kableExtra::column_spec(column = 3, width = "3em") |> 
      kableExtra::column_spec(column = 4, width = "4em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "5em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "8em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "8em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical", 
                                 "C-Cohort, I-Individual",
                                 "RG-Regeneration, RC-Recruitment",
                                 "Y-Yes, N-No",
                                 "D-Daily, M-Monthly"),

                 notation = "alphabet") 
```

## Observed data

Ingrowth data covering a wide range of environmental conditions are hard
to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well ingrowth is
captured. The observations used here are derived from a novel and
unprecedented network of sites in forest reserves that represent the
range of environmental gradients in temperature and precipitation in
Central Europe as compiled in the framework of the EuFoRIa network
[@euforia2019] (Figure \@ref(fig:map)). These forests have been
unmanaged for at least 10 years prior to their designation as reserves,
and most of them provide time series of natural forest dynamics over
multiple decades and up to ca. 80 years. The census periods range from 3
to 37 years, with an average of 14 years. The data provide information
at the tree level, thus allowing for sequential comparison such as
individual-tree ingrowth and death between the measurements.

We selected 200 sites from this network as the benchmarking dataset for
the simulation. The selection of sites is representative of the
environmental variation of the data. This was achieved by applying
k-means clustering to define 200 clusters of plots from the original set
of 869 plots along the environmental dimensions of temperature, climatic
water balance, soil quality and exposition (i.e., slope and aspect).
Each of these sites featured at least two consecutive measurements.
Ingrowth thresholds for the sites differed between diameters of 0 and 10
cm. For our study, we defined two datasets where one included 165 sites
with a diameter threshold of 7 cm or lower, and the other included
another 35 sites with diameter thresholds between 7 and 10 cm.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2022 to provide ingrowth rates
per ha and per decade. The observations used in this study featured
30,900 established trees. Ingrowth rates per site, sample and
decade ranged from zero to 1246 trees per ha, with a mean of 56 trees
per ha. Adult species composition was also available at the plots. For
more details on this unique dataset and the detailed establishment
information, cf. @käber2022.

## Simulation protocol

The overarching goal of the experiments was to assess the tree
establishment as it arises from empirical data against its
representation in a wide range of models of forest dynamics. We define
the tree establishment as the passing of a breast height diameter
threshold of 7 or 10 cm, respectively (synonym: ingrowth). To this end,
each modeling group was provided with a detailed protocol
[@díaz-yáñez2022] with instructions for how to perform the simulations,
providing input variables on climate and soil conditions, and the list
of expected output variables. Neither were further site information
(except for the data specified below) nor any data on tree ingrowth or
forest stand features provided prior to the simulation. That is, the
models were run in "blind flight" mode.

The input variables were collected from different data sources and
aggregated to be adapted to the needs of the different models. Time
series of climatic variables from 1981 to 2018 were provided in hourly
[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
required for some models, such as relative humidity or vapor pressure
deficit, were calculated from these variables. The final instructions
for using the climate data was different in each model, based on the
approach that best suited the model (Table \@ref(tab:tableModels)). Soil
quality data were provided as continuous values between 1 and 5
[Soilgrids dataset, Hengl et al. -@hengl2017]. The protocol also
provided the elevation, slope and aspect for each of the 200 sites
[@asterscienceteam2019], but no other spatial information such as
coordinates, with the exception of iLand and aDGVM2, which required
blurred coordinates to derive highly detailed soil data.

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include a background disturbance to increase
chances of shade-intolerant species to establish. The simulations were
set up to sample species-specific ingrowth rates per decade and per ha
in the equilibrium state of the model, typically entailing a "spin-up"
run (as we did not provide any forest data). The modeling teams decided
on the simulated area and how they derived these samples. The exact
length of the simulation was also decided by the modeling teams (Table
\@ref(tab:tableModels)). Further details on how each modeling team
prepared the simulations and the outputs are available in Supplementary
Material 1.

The simulations were run in the absence of management to a
simulated equilibrium ("Potential Natural Vegetation") with the current
climate, as the ultimate goal was to evaluate tree establishment under
comparable and near equilibrium conditions. This entails the assumption
that the observations from the forest reserves (1) reflect no traces of
forest management and (2) there is an equilibrium between forest
dynamics and climate. While the former might be starting to be visible
in many of the EuFoRIA reserves, the latter may be more debatable.
However, in the absence of detailed data on the history of each plot in
the EuFoRIa network, some broad assumptions had to be made. Both the
width of the regeneration niche (i.e., in environmental space) as well
as the intensity of the ingrowth process (i.e., the number of ingrowth
trees per area and per unit of time) were of interest.

The simulations were run for mixed species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which ingrowth data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
sylvestris L.* The same set of species was used at all 200 sites. Two
models included their standard set of species for the simulations, which
is much larger (i.e., ForClim 1, ForClim 11, TreeMig). In three models,
fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS)
(Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea
abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
parameterised. XComp simulations did not consider *Fraxinus excelsior*
due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
pseudoplatanus* and *Alnus glutinosa* are not parameterized and
therefore these species could not be included in the simulation.
Finally, the model aDGVM2 does not simulate individual species. Rather,
community assembly processes and trait filtering generate plant
communities that are adjusted to the biotic and abiotic conditions, and
the simulated plants can be classified into ecological strategies based
on their trait values in a post-processing step.

Each of the models reported the ingrowth number by sampling 200 times in
a 10-year interval for each species and per ha for each of the 200
sites. Multiple samples per site were used to better understand the
simulated variation within each site. This was done using different
strategies, depending on the model, including (1) sampling simulated
data from the same 1 ha plot in the equilibrium over time, (2) sampling
several 1 ha plots from the simulated forest at one specific point in
time (in the equilibrium) or (3) a combination of (1) and (2) . This
resulted in 880,000 observations per model that simulated the 11 species
included in the protocol (200 sites, 200 samples per site, 11 species
and two diameter thresholds). For the models that simulated additional
species, their ingrowth rates were aggregated as "others".

Two models did not provide results from all the simulated samples or
sites to avoid unrealistic results, as follows. In the model 4C, a
threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
unrealistic stand basal area data, and therefore not all the samples and
sites were reported. The reason for this is that 4C is not suitable for
long-term simulations without management, due to misrepresentations in
density-dependent mortality processes in long-term simulations and
assumptions of tree geometry that lead to unrealistic single tree
dimensions of very old individuals. LPJ-GUESS had 2% of the sample
outputs without tree establishment, and these were considered as zero
stand basal area and zero ingrowth for all the tree species simulated;
one site produced grassland rather than a forest and was not included in
the results.

```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed ingrowth at the 10 cm threshold.", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

## Data analysis

The simulation results were analyzed regarding (1) ingrowth levels
(i.e., ingrowth number per unit time and space), (2) species diversity
of the ingrowth, (3) ingrowth mortality, (4) model performance and model
traits, and (5) ingrowth gradients along regeneration niches. We
evaluated the species diversity across models and in relation to the
observed data by calculating the Shannon index based on the relative
proportion of the species in terms of basal area. The Shannon index was
calculated for the establishment ($H_{R_{n}}$) (Equation
\ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
\ref{eqpiShannon}). The higher the value of the index, the higher the
species diversity at a particular site and sample.

```{=tex}
\begin{equation}
H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
\end{equation}
```
```{=tex}
\begin{equation}
p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
\end{equation}
```
```{=tex}
\begin{equation}
H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where $s$ is the total number of species present and that have a basal
area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
the species $i$ in sample $n$ calculated as the ingrowth basal area
($r_{BA_{n,i}}$) for that species $i$ relative to the total recruited
basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
proportion of the species $i$ calculated as the basal area of all trees
($BA_{n,i}$) of that species $i$ relative to the total basal area
($TotalR_{BA_{n}}$) of the sample $n$. The species diversity was not
assessed for the model aDGVM2 because it does not simulate individual
species.

Mortality in tree establishment was assessed based on the ratio of
ingrowth between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
to estimate whether the ratio of ingrowth between the 7 and 10 cm
diameter thresholds were above or below the expected theoretical rate
between these two diameters. The Reineke self-thinning rule is usually
calculated from even-aged, single species stands and assumes a fixed
relationship between the number of stems and the quadratic mean diameter
in fully stocked pure stands. The value used in our comparisons was 1.77
(i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm),
calculated using Equation \ref{eqReineke}.

```{=tex}
\begin{equation}
Reineke = \frac{7}{10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process based), (3) the presence or absence of a canopy
feedback for ingrowth, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
using two-sided t-tests by considering each of these model features
regarding the observed and simulated ingrowth levels and species
diversity, respectively.

Total ingrowth and the regeneration niches of the individual species
were evaluated across the environmental gradients of light availability,
temperature, and soil moisture. Stand basal area (the basal area of all
the trees in each sample including tree establishment) was used as a
proxy of light availability at the forest floor, the annual degree-day
sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
and the climatic water balance as a proxy for soil moisture
[@speich2019]. Ingrowth values were calculated as the mean across the
200 samples per site. The observed data were modeled using a Generalized
Additive Model [@wood2011] with a negative binomial distribution and
restricted maximum likelihood to better understand the relationship
between the environmental gradients and the levels of observed ingrowth,
relative to the simulation results per model.

In order to analyze the regeneration niches across the climatic
gradients, we focused on the five main tree species or genera: *Abies
alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
*Quercus* spp. For these species, we calculated the share in the
ingrowth basal area per site as the mean across the available samples
per site:

```{=tex}
\begin{equation}
\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
\end{equation}
```
where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
is the mean basal area ingrowth of the species $i$ per site $t$ across
the available samples at that site. Furthermore, regarding the share in
the ingrowth basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
categorized this as zero when the ingrowth basal area of that species
($\bar{r}_{BA_{t,i}}$) was zero but also when both the ingrowth basal
area of that species ($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal
area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.

# Results

## Ingrowth levels

Simulated ingrowth levels varied strongly across the 15 models and
typically did not match the levels found in the forest reserve data
(Figure \@ref(fig:over)). Ingrowth was overestimated in most models for
both the 7 and 10 cm diameter thresholds, with the exception of the
empirical stand model SIBYLA, the landscape model Landis-II and the
global model aDGVM2, which estimated ingrowth levels at the lower end of
the plausibility interval of the observed data (Figure \@ref(fig:over)).
The models with the largest overestimation were the stand model PICUS
and the landscape model TreeMig. For most models, the variability of
simulated ingrowth levels across the 200 sites (visible from the
interquartile range in the box plots of Figure \@ref(fig:over)) was
similar to or smaller than observed, with the notable exception of
PICUS, where simulated ingrowth variability was much larger.

```{r over, echo=FALSE, fig.cap="Mean ingrowth levels across all samples per site, plotted for the 200 sites and for each model. The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data. There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
```

Both the observations and the simulated data had no ingrowth in some
samples and at some sites (for details, cf. Table S1). The observed data
had 4% of the samples with no ingrowth. Only three models had a larger
proportion of no ingrowth (4C, Landis-II and aDVM2). Two models always
simulated ingrowth for both the 7 and 10 cm threshold (xComp and
TreeMig), i.e., they did not feature any zero values. The other ten
models had a very low percentage of samples with no ingrowth (0.01% -
2.39%), i.e., they had distinctly fewer occurrences of zeros compared to
the observations.

## Tree species diversity of ingrowth

Most models matched the level of diversity of the observed data quite
well as quantified by the Shannon index (Figure \@ref(fig:H710)). Five
models overestimated ingrowth diversity: ForCEEPS, ForClim 1, PICUS,
TreeMig and LandClim, the latter particularly for the 7 cm diameter
threshold. The model 4C is a special case, as it simulated five species
only, i.e., its diversity values are not directly comparable to those of
the other models, nor to the observations. Only one model, Landis-II,
consistently underestimated ingrowth diversity.

```{r H710, echo=FALSE, fig.cap="Shannon index of tree ingrowth, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively. The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
    knitr::include_graphics(here::here("figures", "H7_10.png"))
```

In most models, there were only small but significant differences in the
species diversity of ingrowth between 7 and 10 cm across sites (Tables
S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained
the ingrowth diversity between the 7 and 10 cm thresholds (i.e., the
differences between them were not significant, see Table S2), and the
same was evident from the observed data.

For both the observed and the simulated data, and both ingrowth
thresholds, species diversity in the ingrowth compared to stand-level
diversity did not have clear patterns across the models (Figures S1 and
S2). Most models captured reasonably well or overpredicted species
diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left
and center; cf. Figure S3), but a characteristic feature was that
simulations had a much lower variability of diversity compared to
observations. Models overpredicting species diversity included ForClim1,
ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A,
left). Several models underpredicted species diversity, i.e., LPJ-GUESS,
ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).

The majority of the models overpredicted ingrowth diversity; as already
observed in figure \@ref(fig:H710), only two of them underestimated
ingrowth diversity (4C and Landis-II) (Figure
\@ref(fig:combinedRichness)B and Figure S4). Both the models with
feedback and those without feedback from the adult trees to ingrowth
(via seed production) overestimate, underestimated and captured
reasonably well ingrowth species diversity compared to the observations.
Again, simulated diversity was considerably less variable than observed
diversity in the ingrowth.

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data. Each plot shows one pattern represented for one examplary model of each category (overperdicted, intermediate and underpredicted). The full data with the models grouping are shown in Figures S3 and S4. n indicates the number of models falling in each group. A) Examples for the three trends across models for species diversity at the stand level. B) Examples for the three trends across models for ingrowth at the 7 cm threshold.", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

## Mortality in tree establishment

There were strong differences among the models regarding the mortality
rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
models featured a mortality rate significantly larger than the observed
data, thus at least partially compensating for the general
overestimation of ingrowth levels (Figure \@ref(fig:over) and Table S3);
six models did not feature a significant difference compared to the
observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
Four models (xComp, PICUS, LandClim and TreeMig) featured very high
mortality rates (i.e., well above the Reineke self-thinning line), which
compensated for the strong initial overestimation of ingrowth (cf.
Figure \@ref(fig:over)). Interestingly, two models that underestimated
overall ingrowth levels (SIBYLA and Landis-II; cf. Figure
\@ref(fig:over)) featured mortality rates that were close to but still
above the self-thinning line (Figure \@ref(fig:mortality11)).

```{r mortality11, echo=FALSE, fig.cap="Ratio of tree ingrowth rates between the 7 and 10 cm thresholds. Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree ingrowth between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
knitr::include_graphics(here::here("figures", "ratio7_10.png"))
```

The fact that some models (and observed data) featured lower mortality
than expected by Reineke's self-thinning rule may point to facilitation,
or simply a higher mortality rate before the trees had reached 7 cm
diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
whose mortality was well above the self-thinning line, indicates that
they feature higher ingrowth mortality in mixed-species forests compared
to even-aged, single-species forests. Some models yielded a mortality
rate of ingrowth that is broadly compatible with the self-thinning rule
(e.g., 4C and LandClim).

The three models with the largest overestimation of the proportion of
ingrowth at the 7 cm threshold featured the highest mortality between
the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
models that overestimated ingrowth had mortalities lower than the
expected self-thinning ratio, with the exception of LandClim (Figure
\@ref(fig:mortality13)B). It is noteworthy that most of the models that
featured a low ratio of ingrowth between 7 and 10 cm (i.e., ForCEEPS,
ForClim 1, ForClim 11 and iLand) had only a small overestimation of
ingrowth at the 7 cm threshold (Figure \@ref(fig:mortality11)). aDGVM2
was the only model that underestimated ingrowth at the 7 cm threshold
while featuring a low mortality between the two thresholds (Figure
\@ref(fig:mortality13)C), and only a few models were close to the
correct ingrowth levels at 7 cm while also being close to the Reineke
line (SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as
well as LandClim).

```{r mortality13, echo=FALSE, fig.cap="Ratio of tree ingrowth between the 7 and 10 cm thresholds (ingrowth 7 cm / ingrowth 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean ingrowth per model across sites and samples. The horizontal dashed lines show a ratio equal to 1, indicating no decrease of ingrowth between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions. The vertical line indicates zero overestimation.", out.width='80%'}
knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
```

Lastly, some models that had a low number of individuals at the 7 cm
threshold also had a smaller decrease of ingrowth between 7 and 10 cm
(ratio 7-10 cm) (Figure S5). In these models, competition and
self-thinning are either not pronounced or must have occurred before the
species had reached 7 cm. However, this pattern was not consistent
across models (Figure S5). It is noteworthy that there is a relationship
between the ingrowth levels at 7 cm and the mortality rate (Table S4):
most of the models showed a positive effect (higher ingrowth at 7 cm is
coupled to an increase of the mortality between 7 and 10 cm). While some
models showed a negative effect, their linear trend was not pronounced
and not always significant regarding the slope.

## Model performance and model traits

There was no significant relationship between the mean complexity of the
establishment module (Table \@ref(tab:tableModels)) and the
overestimation proportion at a diameter threshold of 7 cm (Figure S6).
All the models except 4C had significant differences between the
observed and simulated mean ingrowth values. Therefore it was not
possible to differentiate in terms of performance between models type
(empirical or process based), or their scales (stand, landscape or
global) (Table S5). The values of ingrowth species diversity simulated
by each model were also assessed against the observed data to see if
there was a difference between models that include a feedback compared
to those that did not; all the models except ForCEEPs(f) and FORMIND had
significant differences (Table S6). It is noteworthy that for the model
ForCEEPS the pattern changed from the overprediction of species
diversity in the ingrowth to a diversity level that is closer to
observations when the recruitment module included feedback
(ForCEEPs(f)).

## Ingrowth gradients on regeneration niches

When evaluating ingrowth levels along key gradients of light
availability (basal area), temperature (degree-days) and soil moisture
(climatic water balance), distinct features emerge: the models
reproduced the effect of basal area in both its magnitude and patterns
(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
than the effects of the climatic gradients, where they featured varying
patterns.

In the observed data, ingrowth levels decreased clearly with increasing
stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a). This trend
was captured by the Landis-II model only, albeit at much lower values of
basal area than in the observations. The other models featured
distinctly different trends, such as (1) an increase in ingrowth levels
with increasing total basal area followed by the absence of ingrowth at
high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and
LPJ-GUESS), (2) almost constant ingrowth levels with basal area (ForClim
11, xComp, LandClim and aDGVM2), or (3) an increase of ingrowth up to a
certain value of stand basal area followed by a decrease at even higher
values of stand basal area, with model-specific thresholds (4C, ForClim
1, PICUS, SIBYLA and iLand).

The observed ingrowth did not change much across the climatic water
balance gradient, and showed a slight increase with temperature. The
models did not match the observed ingrowth patterns across the gradients
of soil moisture and temperature (Figure
\@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a
clear pattern across models. Regarding the soil moisture gradient, a
group of models featured decreasing ingrowth with increasing soil
moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
TreeMig and iLand), while a few models showed the opposite trend
(ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
all the models that featured a decrease in ingrowth levels with
increasing soil moisture showed the opposite trend (i.e., more ingrowth
with increasing temperature), except for 4C. The other models did not
feature a clear pattern.

```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data. For comparing the magnitude of recruitment across models the same graphs with equal y-axis cf. Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}

   knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
                              here::here("figures", "envTrend7WB.png"),
                              here::here("figures", "envTrend7DDS.png")))
```

Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
sylvestris*) and oak (*Quercus* spp.) were the species observed most
frequently in the EuFoRIa data. Most models captured well the share of
basal area in the ingrowth of these five species compared to the
observed data along the two major climatic gradients (Figure
\@ref(fig:diff), cf. Figure S8). However, the models differed
considerably in the way in which they simulated these climatic niches.

First, some models overestimated ingrowth across the entire gradients
for at least one of the five main tree species (Figure \@ref(fig:diff),
e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models
overestimated ingrowth for more than one species for a large part of the
climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red
colors in Figure \@ref(fig:diff)). These trends were consistent for both
ingrowth thresholds, although the exact changes in the share of the
ingrowth basal area were different (Figure S8).

Second, there was some congruence in the simulation results by species
across the models. The share in the ingrowth of Silver fir (*Abies
alba*) was captured well across the gradient by most models except for
xComp, PICUS, and LandClim, which overestimated the ingrowth share of
this species. There were also some peculiarities evident for some
models, such as 4C that did not simulate *Abies alba*. The patterns
across models for *Fagus sylvatica* ingrowth were more complex, as its
ingrowth was underestimated by many models across the environmental
gradients featuring negligible ingrowth at most sites, while others
overestimate it in different parts of the environmental space. Most
models overestimated the share of *Picea abies* ingrowth in the cold-wet
part of the gradients. *Pinus sylvestris* and *Quercus* spp. represented
a smaller share of the ingrowth in the observed data mostly at warm-dry
sites. Many models, however, erroneously featured ingrowth for these two
species along most of the climatic gradients although at a low basal
area share. Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had
almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
feature any ingrowth of these species at all, close to the patterns in
the observed data.

Lastly, there is no model that performed well across the five species
regarding the species-specific ingrowth levels in the environmental
space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
tended to consistently overestimate the ingrowth share of all five
species, but most models overestimated the ingrowth share of some
species while underestimating the share of others (e.g., ForCEEPS,
LPJ-GUESS).

```{r diff, echo=FALSE, fig.cap="Ingrowth levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm ingrowth threshold. The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient. The sizes of the circles represent the ratio between the ingrowth basal area of the species and the total ingrowth basal area for all species. The absence of a circle indicates a zero basal area share in the ingrowth, or the absence of ingrowth altogether. The color gradient (for the models only) shows the difference between the simulated and observed ratio of ingrowth basal area of the species and the total ingrowth basal area for all the species.", out.width='100%', out.height= '85%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
```

# Discussion

Tree establishment is a fundamental process in forest dynamics, in
particular under changing conditions such as those caused by climate
change. Correctly capturing tree ingrowth in dynamic models is
fundamental to, for example, evaluate post-disturbance dynamics and
potential long-term recovery trajectories as it will define the forest
state in the first decades [@seidl2022]. If this initial phase is not
captured well, we can not properly assess aspects such as forest
resilience or the timing, magnitude and progression of carbon
sequestration. In this study, for the first time, projections of tree
establishment from multiple models of forest dynamics were confronted
with a unique dataset from unmanaged forest reserves across a large
environmental gradient in Europe.

The observed data that we used [@käber2022] are exceptional,
particularly with respect to the number of records captured (number of
sites and repeated measurements), which is essential for a highly
'noisy' process such as tree establishment. The use of this dataset for
model benchmarking provided novel insights on the ability of
state-of-the-art models to accurately simulate ingrowth levels, species
composition, and mortality in an early stage of tree life, i.e., between
a tree diameter of 7 and 10 cm. Overall, by adopting this approach a
much broader understanding resulted than if we had used these data for
model calibration: it is primarily from the shortcomings of the models
that we can gain ecological insights [e.g., @trugman2022].

## Ingrowth levels

Most of the models overestimated tree ingrowth levels. This has
potentially far-reaching implications, e.g., regarding biomass (and thus
carbon) turnover, with a potential overestimation of the capacity for
forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
one specific stage, i.e., ingrowth into the 7 and 10 cm diameter
classes. Trees at this size contribute little to carbon sequestration,
and if the excess ingrowth at this stage is compensated -- in the models
-- soon thereafter by higher mortality, overall simulated ingrowth may
still yield trustworthy results at the level of stand structure and
related forest attributes such as basal area or biomass.

From a modeling point of view, excess ingrowth inevitably requires
excess mortality rates in a later stage, either -- as observed for some
models -- between the 7 and 10 cm diameter thresholds, or soon after the
10 cm threshold has been crossed. In any case, correcting at early
stages for the expected forest densities at later stages is equivalent
to compensating for a first error (excess ingrowth) by a second "error"
(excess mortality). It is highly likely that biased projections will
result, because the two errors are unlikely to be perfectly linked and
thus will not always compensate each other. Hence, this structural
problem of most models investigated here is problematic particularly if
the models are to be used under novel conditions such as under climate
change [e.g., @huber2021] or in a decision support context [e.g.,
@thrippleton2021].

Ingrowth levels define the structure and composition of future forests,
but it is equally important to correctly identify areas where ingrowth
is lacking [@rammer2021]. There are multiple constraints to the
regeneration niche of tree species [@price2001], and therefore the
absence of ingrowth is likely to be common [@fortin2007], even over
larger areas such as the one-hectare plots used here. Tree ingrowth data
are characterized by zero-inflation. This was clearly evident from the
EuFoRIa dataset, but some of the models did not produce zeros at all, or
featured a very low proportion of zero data. This substantial difference
may be due to the fact that the simulation results were drawn from
equilibrium forests, whereas in reality many of the forest reserves are
recovering from past management activities and have become denser over
the past decades [e.g., @heiri2009], leading to less ingrowth than in an
equilibrium situation.

## Species composition of ingrowth

Correctly capturing species composition in tree establishment is
important to assess the future functional diversity of a forest, e.g.,
its sensitivity to drought or resilience to disturbances [@redmond2015;
@seidl2022]. In the simulations, overall species diversity levels in the
ingrowth were well within the observed range for half of the models (7
out of 14). Thus, while most of the models are facing difficulties to
quantitatively match ingrowth levels (cf. above), their performance is
better when it comes to matching the composition of ingrowth as a
function of abiotic and biotic conditions. Most models maintained or
even decreased ingrowth diversity between the 7 and 10 cm thresholds,
and the same is visible from the empirical data, but the differences
were not significant. Diversity variations may be explained by the way
the individual models consider establishment processes
[@konig_tree_2022], e.g., without considering the feedback from adults,
ingrowth can be linked to either seed rain from dispersal or a
seed/seedling bank; neither process is incorporated across many models
studied here, however.

The species diversity of the entire stand was captured better by the
models than the diversity of tree ingrowth, and only a few models
overpredicted stand-level diversity. Defining ingrowth composition as
being proportional to adult abundance, regardless of productivity, might
be a simple and conservative assumption to safely maintain relative
species abundances [@hanbury-brown2022], although this may be
simplistic. Furthermore, based on the simulation results, there is no
evidence that models with feedback from the canopy captured better the
species diversity of the ingrowth than those without feedback, with the
exception of ForCEEPS that significantly improved ingrowth species
diversity with the model version that included feedback (ForCEEPS(f)).
The similar performance of models with and without feedback is likely
because the models put more weight on the regeneration niche arising
from abiotic and biotic filters, than from the habitat niche of the
adult trees [@grubb1977].

## Ingrowth mortality

There are several factors that lead to mortality during the
establishment phase, such as competition [@casper1997] and multiple
abiotic factors [@schmid2021; @cunningham2006]. As mentioned above, it
is reasonable to expect that models that overestimate ingrowth may have
a particularly high mortality between the 7 and 10 cm thresholds.
However, this was not consistently evident from the simulations. This
implies that these models must have an excess mortality in later stages,
if we assume that all models were able to capture the structure and
composition of the adult stands along the EuFoRIa gradient; this however
was not tested here.

Higher mortality towards the adult phase has important implications for
forest dynamics and the goods and services provided by forests. On the
one hand, mortality in later stages may erroneously enhance the share of
less shade tolerant species in the models [@klopcic2015], therefore
shifting the species composition. Unrealistic high stem densities for a
longer period of time may overestimate the role of tree regrowth in
carbon sink dynamics [@pugh2019]. Given our set of simulations and
analyses, we cannot conclusively assess what is happening in the models,
and further studies are required that focus on a wider range of tree
sizes and the fate of tree establishment along such a size continuum.

## Model performance and model traits

In spite of the critical considerations above, it is remarkable that
most models did not deviate exceedingly with respect to simulated
ingrowth levels -- after all, few if any of the models' establishment
routines are sufficiently constrained by data, with the exception of the
empirical models xComp and SIBYLA. It is noteworthy that in spite of
their empirical basis, these two models did not match empirical ingrowth
levels, in a similar magnitude as the other models and even in opposite
directions (SIBYLA: underestimation; xComp: overestimation). It appears
that using empirical data of limited geographical scope and constrained
to managed forests, as done in these models, leads to extrapolation
problems already under current climatic conditions. For the other
models, whose establishment modules are not strongly constrained by
empirical data, the degrees of freedom that are available for modeling
ingrowth [@konig_tree_2022] could have implied that model performance
would be much worse than what we found.

Our study showed that increasing complexity in the establishment modules
did not represent an improvement in the projections of ingrowth levels,
species composition or mortality at early tree stages, as there was no
significant relationship between model performance and model complexity.
Increasing complexity in establishment modules has been motivated by
better process understanding and enhancing of model accuracy
[@bugmann2022; @bugmann2019]. However, more complex models do not
necessarily lead to better projections and rather to a reduced
transparency and predictive power [@franklin2020]. Thus, the question of
what level of detail is appropriate and parsimonious for modeling tree
establishment processes remains open [@konig_tree_2022; @bugmann2022].

## Ingrowth gradients on regeneration niches

Competition for light as a strong filter for tree establishment has been
widely documented [@messier1999; @collet2006; @berdanier2016], but the
models examined here did not reproduce this expectation. However, it is
difficult to measure light availability at large spatial and temporal
scales. We used total stand basal area as a proxy for light availability
[cf. @schmid2021]. However we could not consider light availability
restrictions caused by ground vegetation which might be an important
filter for some models [@woltjer2008]. We found pronounced differences
in the stand basal area ranges simulated by the models, but also between
the models and the observations. This made it impossible to evaluate the
establishment for the extremes of the stand density ranges in some
models. For example, ingrowth levels at low stand densities are relevant
to assess how well forests are recovering e.g. after gap creation due to
disturbance [@seidl2022; @grubb1977]. At the level of the simulated
one-hectare samples, average basal area is typically not very low as
long as no larger disturbances occur, which was explicitly excluded in
our simulation protocol. Towards the other end of the spectrum, i.e.,
with increasing stand basal area, it would be reasonable to expect that
the ingrowth of the different species would become sparser and drop out
entirely under low-light conditions [@klopcic2012; @zell2019]. However,
few models showed this trend, thus indicating that the relation between
ingrowth and light availability is not yet captured correctly in most
models. Yet, several of the models that did feature an increase of
ingrowth with increasing basal area include a feedback between seed
production of mature trees and ingrowth. Thus, it seems that higher
light competition does not sufficiently compensate for increased seed
availability with higher basal area due to more mature trees.

There were pronounced differences in how the main tree species were
represented by the models along the environmental gradients, in
particular the dominant species *Abies alba*, *Fagus sylvatica* and
*Picea abies*. The ingrowth levels were sampled from simulations in the
equilibrium, and in this state it is expected that non-climax species
such as *Pinus sylvestris* or *Quercus* would be of minor importance, or
absent [@klopcic2015]. Most models captured this low abundance, which is
also found in the empirical data. Thus, although the broad patterns are
matched by many models, improvements in the quantification of the
regeneration niche of the species are needed, but this cannot be done in
the absence of robust datasets across multiple stand development stages.

## Methodological considerations

The EuFoRIa data as used here are unique and well suited to better
understand tree establishment. However, three aspects of these data may
also represent considerable limitations. First, we made a comparison of
tree ingrowth in an equilibrium state, but we cannot assess how close
the forests included in the EuFoRIa dataset were to such a state. The
data were collected in forest reserves where no management has taken
place for long periods of time. This makes our assumption of an
equilibrium between forest properties and environmental drivers more
reasonable than it might appear at first sight. In an analysis of
primeval forests at demographic equilibrium, @brzeziecki2021 found
higher ingrowth rates than the ones observed in the EuFoRia data. Thus,
the overestimation of ingrowth rates by the models may not be so
problematic. Second, the data were collected for ingrowth above a 7 cm
threshold, thus limiting the assessment of tree establishment to a
holistic viewpoint. This constituted a hard limit based on which we can
understand only some aspects of tree establishment, which in its
entirety often comprises a rather long period since seed production
[@price2001]. In reality, many environmental constraints are acting on
young trees [@käber2021] that we were unable to assess. Yet, ingrowth
data with lower thresholds are simply not available in a harmonized
manner across large environmental gradients. Third, the empirical data
were collected from rather small plots, while we sampled simulated
ingrowth levels from 1 ha areas, which may lead to an incorrect
representation of space. Even though this strategy is not ideal, it
represents a common challenge when harmonizing diverse data sources
originating from varied sampling strategies [@portier2022]. It would
have been extremely challenging for such a variety of models to follow a
protocol where the spatial sampling size was different at each of the
200 sites, and it would have introduced additional uncertainty in the
results. We must not ignore that the observed data represent just a
snapshot of an extremely complex and stochastic reality. This may at
least partly explain e.g., the lack of clear patterns in the simulation
results along the gradients of climate water balance and degree-days.

The design of our sampling protocol did not include spatial aspects such
as seed dispersal or detailed soil data. Although, we considered a wide
range of models of forest dynamics, from stand to global scales, the
simulation setup was limited to 1 ha ingrowth values. This lack of
spatial considerations is appropriate for stand-scale models, but it
potentially puts landscape models at a disadvantage, as they have been
built to be accurate at the landscape level. Without spatial context, we
are limiting ingrowth to the seed influx from the stand itself, unless
the model has a background seed input. However, global models should not
be at a disadvantage due to the limited spatial considerations [but cf.
Snell et al. -@snell2018; or Lehsten et al. -@lehsten2019], as they are
usually lacking dispersal between cells and are based on a strong
abstraction of horizontal space [@hanbury-brown2022]. Lastly, detailed
data on soil conditions were not available from the observed data, and
independent, admittedly rather coarse data for soil and climatic water
balance were used instead. It is noteworthy that several models
represent drought based on more detailed indicators based e.g., on soil
water holding capacity, which had to be derived from a rough soil
quality measure. This may at least partially explain the poor
performance of the models against the drought axis (i.e., climatic water
balance).

## Research recommendations

Based on the results of this study, it is clear that models of forest
dynamics need a focus on tree establishment modules to make them more
robust regarding growth and mortality. It remains uncertain what level
of detail is required to model establishment processes, and this must be
addressed in future research. We recommend that the improvement of the
establishment modules is implemented as additional features that can be
traced back, as done here for the variants of ForClim and ForCEEPS, and
that model complexity and structure must always be connected with
modeling objectives [@albrich2020]. If it should be necessary to include
more detail in the establishment models, this will come with higher
parameterisation efforts. This will most likely lead to lower
generalization because the required data will have to be collected from
specific locations, as currently there is no general establishment
dataset available.

Therefore, we recommend that more effort should be invested into
collecting harmonized datasets on tree establishment. We emphasize that
datasets like EuFoRIa used here are invaluable and should be further
expanded in both their spatial extent (e.g., towards boreal and
Mediterranean conditions) as well as in time (e.g., continuing the
monitoring into the future). Such data will allow for a better
evaluation of forest models and help to reduce the uncertainty in their
projections, which is crucial when they are used as tools for predicting
e.g. the impacts of anthropogenic climate change.

In the present study, we have considered tree establishment in the
equilibrium state only. It is equally important to understand how these
models project tree establishment after changes in forest structure by
disturbances [@seidl2022], or under different management strategies
[@lindner2000]. However, this will require an entirely different set of
observed data, and potentially not all models of forest dynamics would
be able to assess the relationship of these aspects on tree
establishment, e.g., due to the lack of disturbance or appropriate
management modules.

We recommend to further investigate the implications of the current
modeling strategies for tree establishment and, ultimately, simulated
forest stand structure. This applies particularly to the erroneous
patterns of excess tree establishment and later excess mortality, by
focusing on a wider range of tree sizes and the related ingrowth
dynamics. We also recommend, especially for the landscape-level models,
the inclusion of explicit spatial considerations regarding tree ingrowth
[@beckage2003]; this, however, is facing a serious challenge regarding
appropriate datasets, as inventory data have wide coverage, but by
definition do not allow for the assessment of spatial interaction
effects.

Exercises like the one presented here, where the models are operated in
"blind flight" mode, i.e. without the possibility of being tuned towards
capturing the expected patterns, should be repeated. Such benchmarking
exercises should next focus on aspects such as specific model traits and
the ecological formulations of particular (sub-)processes, to better
understand the implications of the assumptions on which the models are
based. Furthermore, the inclusion of a wide range of models with
different scales, type, populations structures, tree establishment
modules and complexity will ensure a wider benefit to the whole modeling
community.

## Conclusions

Models of forest dynamics are important tools in science and decision
support. The formulation of tree establishment has strong implications
for simulated forest properties. The 15 models and variants used here
are facing similar challenges in their representation of tree
establishment: they generally overestimate tree establishment levels,
and the simulated establishment niche is not always captured accurately
as a function of biotic (light) and abiotic (temperature and moisture)
factors.

However, most models properly capture the diversity of the initial tree
community, but differences between model formulations, e.g., the
presence or absence of feedback from the adult trees, did not have a
strong effect for capturing the species composition of ingrowth.

Regarding mortality in the early phase of tree life, many models that
feature a particularly high overestimation of ingrowth levels are
compensating for this by a larger tree mortality. Often, this
compensation is not sufficient to reduce the high ingrowth levels to
realistic values. Overall, there is no clear mortality pattern across
all models.

When capturing tree establishment, the specific design decisions taken
in the development of any model are more important for its behavior
(accuracy) than scale (stand, landscape, global), modelling approach
(empirical vs. process-based), and complexity. Having both empirical and
process-based models in our set, the empirically-based models could have
been expected to have a better performance, as they were calibrated with
inventory data, but this was not the case. Similarly, higher model
complexity does not represent an improvement for capturing tree
establishment.

Even though the establishment routines of most of the models
investigated here have never been sufficiently constrained by robust
data, the ingrowth projections are not overly off. This indicates that a
lot can be gained by a focus on the modeling of establishment processes.
The representation of forest dynamics in these models would become much
more robust particularly in the face of climate change and
post-disturbance dynamics, thus strongly reducing the uncertainty in
long-term projections of future forest dynamics.

# Reproducibility and data availability statement

This study complies with the basic rules of reproducibility and has made
available the protocol of the study, the data and all the software that
has led to the findings presented in this paper. The protocol is
available here @díaz-yáñez2022 . The code and data that support the
findings of this study are openly available at zenodo (*link to be
added*) and github (*link to be added*).

# Acknowledgements

This paper arises from a workshop held in Davos (Switzerland) in June
2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
Funding for the workshop was provided by COST Action CA19139 PROCLIAS
(PROcess-based models for CLimate Impact Attribution across Sectors),
supported by COST (European Cooperation in Science and
Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich
(Research Grant ETH-35 18-1). ODY was funded by the Swiss National
Science Foundation project 200020_188882 / 1 . KB received support from
the European Research Council under the European Union's Horizon 2020
research and innovation program (Grant Agreement 101001905). JB
acknowledges funding from the long-term research development project no.
RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
financial support from I-Maestro (Innovative forest management
strategies for a resilient bioeconomy under climate change and
disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
project REINFORCE (Grant No. 891671). KM received funding from the
Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
no. ITMS2014+313011W580 supported by the Integrated Infrastructure
Operational Programme funded by the ERDF. CH acknowledges financial
support from Waldfonds WASIM2100 project.

# References

```{=html}
<!--The relationship was first described by
[@key] for single citation
[@key1; @key2] multiple citation can be separated by semi-colon
[-@key] in order to suppress author name, and just display the year
[see @key1 p 12; also this ref @key2] is alos a valid syntax
-->
```