### 2023-04-28 manuscript/manuscriptRW.Rmd
changes from Harald

---
title: "Tree regeneration in models of forest dynamics: a key priority for further research"
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Friedrich Bohn,
- Kristin H. Braziunas,
- Josef	Brůna,
- Rico Fischer,
- Samuel M. Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred J. Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Ján Merganič,
- Katarina Merganičová,
- Tobias Mette,
- Marco Mina,
- Xavier Morin,
- Werner Rammer,
- Christopher P.O. Reyer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
    extra_dependencies: "subfig"
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2007, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016, @langan2017, @scheiter2013 
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

```{r, echo=FALSE}
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries = FALSE)
  library(bibtex)
  options(tinytex.verbose = TRUE)
```

# Abstract

Tree regeneration is a key process in forest dynamics, particularly in
the context of forest resilience and climate change. Models are pivotal
for assessing long-term forest dynamics, and they have been in use for
more than 50 years. However, an assessment of their ability to
accurately represent tree regeneration is lacking. We assess how well
current models capture the overall abundance, species composition, and
mortality of tree regeneration.

Using 15 models built to capture long-term forest dynamics at the stand,
landscape, and global levels, we simulate tree regeneration at 200 sites
representing large environmental gradients across Central Europe. The
results are evaluated against comprehensive data from unmanaged forests.

Most of the models overestimate regeneration levels, which is only
compensated by some models with high simulated mortality rates in the
early stages of individual trees dynamics. Simulated species diversity
of regeneration matches the observed ranges. Models simulating higher
species diversity at the stand level do not feature higher regeneration
diversity. The effect of light availability on regeneration levels is
captured better than the effect of temperature and soil moisture, but
patterns are not consistent across models. Increasing complexity in the
tree regeneration modules of the models is not related to higher
accuracy of simulated tree regeneration. Furthermore, individual model
design is more important than scale (stand, landscape, global) and
approach (empirical, process-based) for accurately capturing tree
regeneration.

Despite considerable mismatches between simulation results and data in
some instances, it is remarkable that most of the models capture the
essential features of the highly complex process of tree regeneration,
while not having been parameterized with such data. We conclude that
much can be gained by evaluating and refining model regeneration
processes. This has the potential to render long-term projections of
forest dynamics under changing environmental conditions that are much
more robust.

\newpage

# Introduction

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbance impacts and changing societal
demands on forest ecosystem services, it is becoming ever more important
to understand how forest structure, composition and function will
change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics were developed over the past decades considering the
impacts of climate [@bugmann2022]. From these studies, it is evident
that we have a reasonably good understanding how to model tree growth
[@vanclay1997; @bugmann1996], and substantial efforts have been
dedicated to improving the representation of tree mortality
[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
less studied process that is often represented rather coarsely in models
[@walck2011; @leishman1992; @price2001]. Which is a clear research gap
in the context of climate-induced forest disturbances and forest
resilience.

Tree regeneration arises from multiple processes such as pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree regeneration is
challenging. Currently, tree regeneration processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
of a few simple environmental filters, as done in most forest "gap"
models [@shuman2011] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994]. Overall, models are needed to
(1) synthesize existing empirical data and explore their relationships,
(2) assess future tree regeneration e.g. in the context of global change
scenarios, and (3) identify the most important processes that are
shaping ecological patterns. Given the current strategies that are used
in models of forest dynamics to represent tree regeneration, model
behavior often is prone to problems, such as very high levels of tree
regeneration that necessitate excess mortality at early stages of tree
life to simulate correct stand structure and composition. Also,
inadequately high species diversity in tree regeneration may be
simulated, which is characteristic of "classical" forest gap models
[@gutiérrez2016], at least as long as the simulation set-up comprises a
multitude of species. Some models use calibration against local
canopy-level data to constrain simulated regeneration levels, which is
likely to hamper the general applicability of these models e.g. under
scenarios of climate change. Furthermore, correctly capturing the
species composition in tree regeneration as a function of the presence
of seed trees in the canopy is often a particular challenge, potentially
leading to unrealistic successional drift in the model, which must be
corrected by factors that are hard to parameterize [e.g., @lischke2006].
A related issue is the excessive reduction of species diversity due to
positive feedback effects, such that eventually just single-species
stands remain. This is sometimes corrected by the incorporation of a low
level of seed influx of all species at all times [@schumacher2006] or by
restricting the number of seeds per species in the seedbank
[@lischke2006]. However, simulated species composition is usually
exceedingly sensitive to assumptions about seed availability, whereas
the parameters of such functions are poorly constrained by field data.
Lastly, there is often a problem with insufficient observational
constraints on parameter values for models that start from very small
tree sizes (e.g., 10 cm of height or even seeds) and track tree
development in a process-oriented manner by considering a multitude of
ecological influences, rather than emphasizing tree regeneration into a
larger size class [e.g., @wehrli2006].

Thus, a focus on the modeling of tree regeneration processes is sorely
needed and overdue [@price2001; @walck2011] if we are to make reliable
projections of future forest dynamics, i.e., when the models need to be
operated in extrapolation mode, as well as from a fundamental ecological
point of view for increasing systems understanding. Also in empirical
ecology, tree regeneration has not received a lot of attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail [@miina2006]. Yet, little data are available that covers all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Thus, the available data are fragmented,
which constitutes a major problem for model building [@clark1999]. On
the one hand, monitoring on permanent plots (such as National Forest
Inventories) often measures tree regeneration [@zell2019], however with
a design that captures the rate after the stand initiation phase
[@hallsby2015]. Data from permanent plots, although available across
regions, are highly heterogeneous, rendering their use difficult in a
modeling context [@käber2021]. On the other hand, targeted studies to
measure tree regeneration on experimental sites are often limited in
spatial extent, species studied, or the sub-set of processes that are
targeted [@berdanier2016; @collet2006].

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree regeneration [@käber2022]. By tree regeneration, we refer to the
passing of trees across a specific diameter threshold ("ingrowth"). We
included models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving this process --- including some that are not incorporated
explicitly in most models, such as deer browsing --- we do not aim for a
detailed statistical evaluation of each model. Instead, we aim to
evaluate the general regeneration patterns and magnitudes simulated by
the models, and benchmark the simulated regeneration niche of multiple
species against empirical data along a wide environmental gradient of
temperature, moisture, and light availability [@grubb1977].

More specifically, we aim to answer the following questions: (1) Are
models of forest dynamics capturing accurately tree regeneration levels,
initial tree species diversity and mortality in the regeneration? (2) Do
model traits explain differences in model performance? (3) How well do
the models capture total regeneration and individual species
regeneration niches across environmental gradients of light
availability, temperature, and soil moisture?

# Material and methods

## Models

Fifteen models of forest dynamics (including two models featuring model
variants) were used to simulate forest dynamics at the stand, landscape
or global scale (Table \@ref(tab:tableModels)). The approaches used for
model construction and their origin differ strongly, with most of the
models featuring a largely "process-based" approach, whereas two models
are based on formulations derived from the statistical analysis of
inventory data (SIBYLA and xComp). Some of the models largely rely on
the original approach underlying forest gap models (e.g., ForClim 1),
whereas some are based on plant ecophysiological processes (e.g.,
FORMIND and iLand). The different approaches underlying the models have
strongly influenced the formulation of tree regeneration processes.

The regeneration models can be differentiated into 'regeneration' and
'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
models* include processes such as flowering and pollination, seed
production, seed dispersal, germination, and seedling growth, which
ultimately lead to the simulated number of established trees.
*Recruitment models*, in contrast, introduce a number of new trees with
certain characteristics such as biomass or diameter, without explicitly
considering earlier development processes. We can further distinguish
models that feature a feedback in the regeneration, i.e., where the
level and species composition of regeneration is influenced by the
existing adult tree community via the production of seeds, seedlings or
saplings, from models that do not contain such a feedback.

The starting point for tree regeneration in the models differs as well,
ranging from seedbank, seed or seedling to sapling (i.e., trees much
larger than 10 cm height, often ca. 2-3 m tall). Models that start from
seed need to include a larger number of ecological processes such as
germination and survival of young seedlings, whereas models that start
from saplings have to aggregate via parameterizations several ecological
processes that are not treated explicitly. This latter approach reduces
model complexity but comes at the cost of blurring process
representation.

In the models used here, the overall complexity in the regeneration
modules varies considerably. Following @bugmann2022, we can classify
seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
TreeMig) as having rather high complexity in their regeneration modules
(mean regeneration formulation complexity across all processes \>0,
Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
used here with two alternative variants of regeneration. ForClim variant
1 [@bugmann1996] is based on a recruitment module that adheres closely
to the concept introduced by @botkin1972, whereas ForClim variant 11 is
adopting a different approach where individual species properties and
their relation to the environment are important [@huber2020]. These two
model variants allow us to evaluate a more process-based and complex
module (variant 11) against a simple module (variant 1), while the rest
of the model structure is identical. Similarly, the two ForCEEPS
variants allow us to isolate the importance of the canopy feedback
(i.e., simulated actual composition and relative abundance of species in
the plot) via seed trees for the quantity and quality (e.g., diversity
and composition) of simulated regeneration, as one variant includes this
feedback whereas the other does not.

```{r tableModels, echo=FALSE,  message=FALSE, warning=TRUE}
library(knitr)
modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))
modelDF$Sampling <- NULL

#footnotes
names(modelDF)[2] <- paste0(names(modelDF)[2], "(a)")
names(modelDF)[3] <- paste0(names(modelDF)[3], "(b)")
names(modelDF)[4] <- paste0(names(modelDF)[4], "(c)")
names(modelDF)[6] <- paste0(names(modelDF)[6], "(d)")
names(modelDF)[11] <- paste0(names(modelDF)[11],"(e)")

kable(modelDF,
      format = "latex", booktabs = TRUE,
      caption = "Models included and their characteristics.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |>
      kableExtra::column_spec(column = 2, width = "3em") |>
      kableExtra::column_spec(column = 3, width = "4em") |>
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "8em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "5em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "8em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical",
                                 "C-Cohort, I-Individual",
                                 "Y-Yes, N-No",
                                 "D-Daily, M-Monthly"),

                 notation = "alphabet")
```

## Observed data

Regeneration data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel and unprecedented network of sites in forest reserves that
represent the range of environmental gradients in temperature and
precipitation in Central Europe as compiled in the framework of the
EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)). These forests
have been unmanaged for at least 10 years prior to their designation as
reserves, and most of them provide time series of natural forest
dynamics over multiple decades and up to ca. 80 years. The census
periods range from 3 to 37 years, with an average of 14 years. The data
provide information at the tree level, thus allowing for the sequential
comparison of processes such as individual-tree regeneration and death
between the measurements.

We selected 200 sites from this network as the benchmarking dataset for
the simulation to be representative of the environmental variation
contained in the data. This was achieved by applying k-means clustering
to define 200 clusters of plots from the original set of 869 plots along
the environmental dimensions of temperature, climatic water balance,
soil quality and exposition (i.e., slope and aspect). Each of these
sites featured at least two consecutive measurements. Regeneration
thresholds for these sites differed between diameters of 0 and 10 cm.
For our study, we defined two datasets where one included 165 sites with
a diameter threshold of 7 cm or lower, and the other included another 35
sites with diameter thresholds between 7 and 10 cm.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2022 to provide regeneration
rates per ha and per decade. The observations used in this study
featured 30,900 newly established trees. Regeneration rates per site,
sample and decade ranged from zero to 1246 trees, with a mean of 56
trees. Adult species composition was also available for each plot. For
more details on this unique dataset and the detailed regeneration
information, cf. @käber2022.

## Simulation protocol

The overarching goal of the experiments was to assess the tree
regeneration as it arises from empirical data against its representation
in a wide range of models of forest dynamics. We define tree
regeneration as the passing of a breast height diameter threshold of 7
or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
group was provided with a detailed protocol [@díaz-yáñez2022] with
instructions how to perform the simulations, providing input variables
on climate and soil conditions, and the list of expected output
variables. Neither were further site information (except for the data
specified below) nor any data on tree regeneration or forest stand
features provided prior to the simulation. That is, the models were run
in "blind flight" mode.

The input variables were collected from different data sources and
aggregated to be adapted to the needs of the different models. Time
series of climatic variables from 1981 to 2018 were provided in hourly
[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
required for some models, such as relative humidity or vapor pressure
deficit, were calculated from these variables. The final instructions
for using the climate data were slightly different in each model, based
on the approach that best suited the model (Table
\@ref(tab:tableModels)). Soil quality data were provided as continuous
values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
The protocol also provided the elevation, slope and aspect for each of
the 200 sites [@asterscienceteam2019], but no other spatial information
such as coordinates, with the exception of iLand and aDGVM2, which
required blurred coordinates to derive highly detailed soil data.

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include a background disturbance to increase
chances of shade-intolerant species to establish. The simulations were
set up to sample species-specific regeneration rates per decade and per
ha in the equilibrium state of the model, typically entailing a
"spin-up" run (as we did not provide any forest data). The modeling
teams decided on the simulated area and how they derived these samples.
The exact length of the simulation was also decided by the modeling
teams (Table \@ref(tab:tableModels)). Further details on how each
modeling team prepared the simulations and the outputs are available in
Supplementary Material 1.

The simulations were run in the absence of management to a simulated
equilibrium ("Potential Natural Vegetation") with the current climate,
as the ultimate goal was to evaluate tree regeneration under comparable
and near-equilibrium conditions. This entails the assumption that (1)
the observations from the forest reserves reflect no traces of forest
management, and (2) there is an equilibrium between forest dynamics and
climate. While the former might be starting to be visible in many of the
EuFoRIA reserves, the latter may be more debatable. However, in the
absence of detailed data on the history of each plot in the EuFoRIa
network, some broad assumptions had to be made. Both the width of the
regeneration niche (i.e., in environmental space) as well as the
intensity of the regeneration process (i.e., the number of ingrowth
trees per area and per unit of time) were of interest.

The simulations were run for mixed-species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which regeneration data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
sylvestris L.* The same set of species was used at all 200 sites. Two
models included their standard set of species for the simulations, which
is much larger (i.e., ForClim 1, ForClim 11, TreeMig). In three models,
fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS)
(Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea
abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
parameterised. xComp simulations did not consider *Fraxinus excelsior*
due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
pseudoplatanus* and *Alnus glutinosa* are not parameterized and
therefore these species could not be included in the simulation.
Finally, the model aDGVM2 does not simulate individual species. Rather,
community assembly processes and trait filtering generate plant
communities that are adjusted to the biotic and abiotic conditions, and
the simulated plants can be classified into ecological strategies based
on their trait values in a post-processing step.

Each of the models reported the regeneration number by sampling 200
times in a 10-year interval for each species and per ha for each of the
200 sites. Multiple samples per site were used to better understand the
simulated variation within each site. This was done using different
strategies, depending on the model, including (1) sampling simulated
data from the same 1 ha plot in the equilibrium over time, (2) sampling
several 1 ha plots from the simulated forest at one specific point in
time (in the equilibrium) or (3) a combination of (1) and (2). This
resulted in 880,000 observations per model that simulated the 11 species
included in the protocol (200 sites, 200 samples per site, 11 species
and two diameter thresholds). For the models that simulated additional
species, their regeneration rates were aggregated as "others".

Two models did not provide results from all the simulated samples or
sites to avoid unrealistic results, as follows. In the model 4C, a
threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
unrealistic stand basal area data, and therefore not all the samples and
sites were reported. The reason for this is that 4C is not suitable for
long-term simulations without management, due to misrepresentations in
density-dependent mortality processes in long-term simulations and
assumptions of tree geometry that lead to unrealistic single tree
dimensions of very old individuals. LPJ-GUESS had 2% of the sample
outputs without tree regeneration, and these were considered as zero
stand basal area and zero regeneration for all the tree species
simulated; one site produced grassland rather than a forest and was not
included in the results.

```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

## Data analysis

The simulation results were analyzed regarding (1) regeneration levels
(i.e., ingrowth number per unit time and space), (2) regeneration
species diversity, (3) regeneration mortality, (4) the relationship
between model performance and model traits, and (5) ingrowth gradients
along the regeneration niches. We evaluated species diversity across the
models and in relation to the observed data by calculating the Shannon
index $H$ based on the relative proportion of the species in terms of
basal area. It was calculated for the regeneration ($H_{R_{n}}$)
(Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$)
(Equation \ref{eqpiShannon}). The higher the value of the index, the
higher the species diversity at a particular site and sample.

```{=tex}
\begin{equation}
H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
\end{equation}
```
```{=tex}
\begin{equation}
p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
\end{equation}
```
```{=tex}
\begin{equation}
H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where $s$ is the total number of species present and that have a basal
area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
species $i$ in sample $n$ calculated as the regeneration basal area
($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
proportion of species $i$ calculated as the basal area of all trees
($BA_{n,i}$) of that species $i$ relative to total basal area
($TotalR_{BA_{n}}$) of the sample $n$. Species diversity was not
assessed for aDGVM2 as this model does not simulate individual species.

Mortality in tree regeneration was assessed based on the ratio of
regeneration between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
to estimate whether the ratio of regeneration between the 7 and 10 cm
diameter thresholds were above or below the expected theoretical rate.
The Reineke self-thinning rule is usually calculated for even-aged,
single species stands and assumes a fixed relationship between the
number of stems and the quadratic mean diameter in fully stocked pure
stands. The value used in our comparisons was 1.77 (i.e., we expect stem
numbers at 7 cm to be 77% higher than at 10 cm), calculated using
Equation \ref{eqReineke}.

```{=tex}
\begin{equation}
Reineke = \frac{7}{10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
using two-sided t-tests by considering each of these model features
regarding the observed and simulated regeneration levels and species
diversity, respectively.

Total regeneration and the regeneration niches of the individual species
were evaluated across the environmental gradients of light availability,
temperature, and soil moisture. Stand basal area (the basal area of all
the trees in each sample including tree regeneration) was used as a
proxy for light availability at the forest floor, the annual degree-day
sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
and the climatic water balance as a proxy for soil moisture
[@speich2019]. Regeneration values were calculated as the mean across
the 200 samples per site. The observed data were modeled using a
Generalized Additive Model [@wood2011] with a negative binomial
distribution and restricted maximum likelihood to better understand the
relationship between the environmental gradients and the levels of
observed regeneration, relative to the simulation results per model.

In order to analyze the regeneration niches across the climatic
gradients, we focused on five common tree species or genera: *Abies
alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
*Quercus* spp. For these species, we calculated the share in the
regeneration basal area per site as the mean across the available
samples per site:

```{=tex}
\begin{equation}
\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
\end{equation}
```
where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
is the mean basal area in the regeneration (subsequently referred to as
"regeneration basal area") of species $i$ per site $t$ across the
available samples at that site. Furthermore, regarding the share in the
regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
categorized this as zero when the regeneration basal area of that
species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the
regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) and the
total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$)
equaled zero.

# Results

## Regeneration levels

Simulated regeneration levels varied strongly across the 15 models and
typically did not match the levels found in the forest reserve data
(Figure \@ref(fig:over)). Regeneration was overestimated in most models
for both the 7 and 10 cm diameter thresholds, with the exception of the
empirical stand model SIBYLA, the landscape model Landis-II and the
global model aDGVM2, which estimated regeneration levels at the lower
end of the plausibility interval of the observed data (Figure
\@ref(fig:over)). The models with the largest overestimation were the
stand model PICUS and the landscape model TreeMig. For most models, the
variability of simulated regeneration levels across the 200 sites
(visible from the interquartile range in the box plots of Figure
\@ref(fig:over)) was similar to or smaller than observed, with the
notable exception of PICUS, where simulated regeneration variability was
much larger.

```{r over, echo=FALSE, fig.cap="Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model. The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data. There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
```

Both the observations and the simulated data had no regeneration in some
samples and at some sites (for details, cf. Table S1). The observed data
had 4% of the samples with no regeneration. Only three models had a
larger proportion of no regeneration (4C, Landis-II and aDVM2). Two
models always simulated regeneration for both the 7 and 10 cm threshold
(xComp and TreeMig), i.e., they did not feature any zero values. The
other ten models had a very low percentage of samples with no
regeneration (0.01% - 2.39%), i.e., they had distinctly fewer
occurrences of zeros compared to the observations.

## Tree species diversity of regeneration

Most models matched the level of diversity of the observed data quite
well (Figure \@ref(fig:H710)). Five models overestimated regeneration
diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
particularly for the 7 cm diameter threshold. The model 4C is a special
case, as it simulated five species only, i.e., its diversity values are
not directly comparable to those of the other models, nor to the
observations. Only one model, Landis-II, consistently underestimated
regeneration diversity.

```{r H710, echo=FALSE, fig.cap="Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively. The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
    knitr::include_graphics(here::here("figures", "H7_10.png"))
```

In most models, there were only small but significant differences in the
species diversity of regeneration between 7 and 10 cm across sites
(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
maintained the regeneration diversity between the 7 and 10 cm thresholds
(i.e., the differences between them were not significant, see Table S2),
and the same was evident from the observed data.

For both the observed and the simulated data, and both regeneration
thresholds, species diversity in the regeneration compared to
stand-level diversity did not feature clear patterns across the models
(Figures S1 and S2). Most models captured reasonably well or
overpredicted species diversity at the stand level (Figure
\@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a
characteristic feature was that the simulations had a much lower
variability of diversity compared to observations. Models overpredicting
species diversity included ForClim1, ForClim 11, FORMIND and TreeMig
(Figure \@ref(fig:combinedRichness)A, left). Several models
underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C
(Figure \@ref(fig:combinedRichness)A, right).

The majority of the models overpredicted regeneration diversity; as
already observed in Figure \@ref(fig:H710), only two of them
underestimated it (4C and Landis-II) (Figure
\@ref(fig:combinedRichness)B and Figure S4). Both the models with
feedback and those without feedback from the adult trees to regeneration
(via seed production) overestimated, underestimated or captured
reasonably well regeneration species diversity compared to the
observations. Again, simulated diversity had considerably lower
variability than observed diversity in the regeneration.

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data. Each plot shows one pattern represented for one examplary model of each category (overperdicted, intermediate and underpredicted). The full data with the models grouping are shown in Figures S3 and S4. n indicates the number of models falling in each group. A) Examples for the three trends across models for species diversity at the stand level. B) Examples for the three trends across models for regeneration at the 7 cm threshold.", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

## Mortality in tree regeneration

There were strong differences among the models regarding the mortality
rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
models featured a mortality rate significantly larger than the observed
data, thus at least partially compensating for the general
overestimation of regeneration levels (Figure \@ref(fig:over) and Table
S3); six models did not feature a significant difference compared to the
observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
Four models (xComp, PICUS, LandClim and TreeMig) featured very high
mortality rates (i.e., well above the Reineke self-thinning line), which
compensated for the strong initial overestimation of regeneration (cf.
Figure \@ref(fig:over)). Interestingly, two models that underestimated
overall regeneration levels (SIBYLA and Landis-II; cf. Figure
\@ref(fig:over)) featured mortality rates that were close to but still
above the self-thinning line (Figure \@ref(fig:mortality11)).

```{r mortality11, echo=FALSE, fig.cap="Ratio of tree regeneration rates between the 7 and 10 cm thresholds. Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
knitr::include_graphics(here::here("figures", "ratio7_10.png"))
```

The fact that some models (and observed data) featured lower mortality
than expected by Reineke's self-thinning rule may point to facilitation,
or simply a higher mortality rate before the trees had reached 7 cm
diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
whose mortality was well above the self-thinning line, indicates that
they feature higher regeneration mortality in mixed-species forests
compared to even-aged, single-species forests. Some models yielded a
mortality rate of regeneration that is broadly compatible with the
self-thinning rule (e.g., 4C and LandClim).

The three models with the largest overestimation of the proportion of
regeneration at the 7 cm threshold featured the highest mortality
between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The
other models that overestimated regeneration had mortalities lower than
the expected self-thinning ratio, with the exception of LandClim (Figure
\@ref(fig:mortality13)B). It is noteworthy that most of the models that
featured a low ratio of regeneration between 7 and 10 cm (i.e.,
ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
overestimation of regeneration at the 7 cm threshold (Figure
\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
regeneration at the 7 cm threshold while featuring a low mortality
between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
few models were close to the correct regeneration levels at 7 cm while
also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
\@ref(fig:mortality13)D; and 4C as well as LandClim).

```{r mortality13, echo=FALSE, fig.cap="Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean regeneration per model across sites and samples. The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions. The vertical line indicates zero overestimation.", out.width='80%'}
knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
```

Lastly, some models that had a low number of individuals at the 7 cm
threshold also had a smaller decrease of regeneration between 7 and 10
cm (ratio 7-10 cm) (Figure S5). In these models, competition and
self-thinning are either not pronounced or must have occurred before the
trees had reached 7 cm. However, this pattern was not consistent across
models (Figure S5). It is noteworthy that there is a relationship
between the regeneration levels at 7 cm and the mortality rate (Table
S4): most of the models showed a positive effect (higher regeneration at
7 cm is coupled to an increase of the mortality between 7 and 10 cm).
While some models showed a negative effect, the linear trend was not
pronounced and not always significant regarding the slope.

## Model performance and model traits

There was no significant relationship between the mean complexity of the
regeneration module (Table \@ref(tab:tableModels)) and the
overestimation proportion at a diameter threshold of 7 cm (Figure S6).
All the models except 4C had significant differences between the
observed and simulated mean regeneration values. Therefore it was not
possible to differentiate in terms of performance between models type
(empirical or process based), or their scales (stand, landscape or
global) (Table S5). The values of regeneration species diversity
simulated by each model were also assessed against the observed data to
see if there was a difference between models that include a feedback
compared to those that did not; all the models except ForCEEPs(f) and
FORMIND had significant differences (Table S6). It is noteworthy that
for the model ForCEEPS the pattern changed from the overprediction of
species diversity in the regeneration to a diversity level that is
closer to observations when the recruitment module included feedback
(ForCEEPs(f)).

## Regeneration gradients on regeneration niches

When evaluating regeneration levels along key gradients of light
availability (basal area), temperature (degree-days) and soil moisture
(climatic water balance), distinct features emerged: the models
reproduced the effect of basal area in both its magnitude and patterns
(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
than the effects of the climatic gradients, where they featured varying
patterns.

In the observed data, regeneration levels decreased clearly with
increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
This trend was captured by the Landis-II model only, albeit at much
lower values of basal area than in the observations. The other models
featured distinctly different trends, such as (1) an increase in
regeneration levels with increasing total basal area followed by the
absence of regeneration at high values of basal area (ForCEEPS,
ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant
regeneration levels with basal area (ForClim 11, xComp, LandClim and
aDGVM2), or (3) an increase of regeneration up to a certain value of
stand basal area followed by a decrease at even higher values of stand
basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
and iLand).

The observed regeneration did not change much across the climatic water
balance gradient, and showed a slight increase with temperature. The
models did not match the observed regeneration patterns across the
gradients of soil moisture and temperature (Figure
\@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a
clear pattern across models. Regarding the soil moisture gradient, a
group of models featured decreasing regeneration with increasing soil
moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
TreeMig and iLand), while a few models showed the opposite trend
(ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
all the models that featured a decrease in regeneration levels with
increasing soil moisture showed the opposite trend (i.e., more
regeneration with increasing temperature), except for 4C. The other
models did not feature a clear pattern.

```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data. For comparing the magnitude of recruitment across models the same graphs with equal y-axis cf. Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}

   knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
                              here::here("figures", "envTrend7WB.png"),
                              here::here("figures", "envTrend7DDS.png")))
```

Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
sylvestris*) and oak (*Quercus* spp.) were the species observed most
frequently in the EuFoRIa data. Most models captured well the share of
basal area in the regeneration of these five species compared to the
observed data along the two major climatic gradients (Figure
\@ref(fig:diff), cf. Figure S8). However, the models differed
considerably in the way in which they simulated these climatic niches.

First, some models overestimated regeneration across the entire
gradients of at least one of the five main tree species (Figure
\@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS),
while other models overestimated regeneration of more than one species
for a large part of the climate gradients, such as xComp, PICUS, SIBYLA
or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)). These trends
were consistent for both regeneration thresholds, although the exact
changes in the share of the regeneration basal area were different
(Figure S8).

Second, there was some congruence in the simulation results by species
across the models. The share in the regeneration of Silver fir (*Abies
alba*) was captured well across the gradient by most models except for
xComp, PICUS, and LandClim, which overestimated the regeneration share
of this species. There were also some peculiarities evident for some
models, such as 4C that did not simulate *Abies alba*. The patterns
across models for *Fagus sylvatica* regeneration were more complex, as
its regeneration was underestimated by many models across the
environmental gradients featuring negligible regeneration at most sites,
while others overestimated it in different parts of the environmental
space. Most models overestimated the share of *Picea abies* regeneration
in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
spp. represented a small share of the regeneration in the observed data
mostly at warm-dry sites. Many models, however, erroneously featured
regeneration for these two species along most of the climatic gradients,
although with a low share of basal area. Five models (4C, FORMIND,
SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus
sylvestris* and *Quercus* spp., or did not feature any regeneration of
these species at all.

Lastly, there is no model that performed well across the five species
regarding the species-specific regeneration levels in the environmental
space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
tended to consistently overestimate the regeneration share of all five
species, but most models overestimated the regeneration share of some
species while underestimating the share of others (e.g., ForCEEPS,
LPJ-GUESS).

```{r diff, echo=FALSE, fig.cap="Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold. The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient. The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species. The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether. The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.", out.width='100%', out.height= '85%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
```

# Discussion

Tree regeneration is a fundamental process in forest dynamics. Correctly
capturing tree regeneration in dynamic models is fundamental to, for
example, evaluate post-disturbance dynamics and potential long-term
recovery trajectories as it will define the forest state in the first
decades [@seidl2022]. If this initial phase is not captured well, we can
not properly assess aspects such as forest resilience or the timing,
magnitude and progression of carbon sequestration. In this study, for
the first time, projections of tree regeneration from multiple models of
forest dynamics were confronted with a unique dataset from unmanaged
forest reserves across a large environmental gradient in Europe.

The EuFoRIa data [@käber2022] are exceptional, particularly with respect
to the number of records (number of sites and repeated measurements),
which is essential for capturing a highly 'noisy' process such as tree
regeneration. The use of this dataset for model benchmarking provided
novel insights on the ability of state-of-the-art models to accurately
simulate regeneration levels, species composition, and mortality in an
early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
Overall, by adopting this approach a much broader understanding resulted
than if we had used these data for model calibration: it is primarily
from the shortcomings of the models that we can gain ecological insights
[cf. @trugman2022].

## Regeneration levels

Most of the models overestimated tree regeneration levels. This has
potentially far-reaching implications, e.g., regarding biomass (and thus
carbon) turnover, with a potential overestimation of the capacity for
forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
one specific stage, i.e., regeneration into the 7 and 10 cm diameter
classes. Trees at this size contribute little to carbon sequestration,
and if the excess regeneration at this stage is compensated -- in the
models -- soon thereafter by higher mortality, overall simulated
regeneration may still yield trustworthy results at the level of stand
structure and related forest attributes such as basal area or biomass.

From a modeling point of view, excess regeneration inevitably requires
excess mortality rates in a later stage, either -- as observed for some
models -- between the 7 and 10 cm diameter thresholds, or soon after the
10 cm threshold has been crossed. In any case, correcting at early
stages for the expected forest densities at later stages is equivalent
to compensating for a first error (excess regeneration) by a second
error (excess mortality). It is highly likely that biased projections
will result, because the two errors are unlikely to be perfectly linked
and thus will not always compensate each other. Hence, this structural
problem of most models investigated here is problematic particularly if
the models are to be used under novel conditions such as under climate
change [e.g., @huber2021] or in a decision support context [e.g.,
@thrippleton2021].

Regeneration levels define the structure and composition of future
forests, but it is equally important to correctly identify areas where
regeneration is lacking [@rammer2021]. There are multiple constraints to
the regeneration niche of tree species [@price2001], and therefore the
absence of regeneration is likely to be common [@fortin2007], even over
larger areas such as the one-hectare plots used here. Tree regeneration
data are characterized by zero-inflation. This was clearly evident from
the EuFoRIa dataset, but some of the models did not produce zeros at
all, or featured a very low proportion of zero data. This substantial
difference may be due to the fact that the simulation results were drawn
from equilibrium forests, whereas in reality many of the forest reserves
are recovering from past management activities and have become denser
over the past decades [e.g., @heiri2009], leading to less regeneration
than in an equilibrium situation.

## Species composition of regeneration

Correctly capturing species composition in tree regeneration is
important to assess the future functional diversity of a forest, e.g.,
its sensitivity to drought or resilience to disturbances [@redmond2015;
@seidl2022]. In the simulations, overall species diversity levels in the
regeneration were well within the observed range for half of the models
(7 out of 14). Thus, while most of the models are facing difficulties to
quantitatively match regeneration levels (cf. above), their performance
is better when it comes to matching the composition of regeneration as a
function of abiotic and biotic conditions. Most models maintained or
even decreased regeneration diversity between the 7 and 10 cm
thresholds, and the same is visible from the empirical data, but the
differences were not significant. Diversity variations may be explained
by the way the individual models consider regeneration processes
[@konig_tree_2022], e.g., without considering the feedback from adults,
regeneration can be linked to either seed rain from dispersal or a
seed/seedling bank; neither process is incorporated across many models
studied here, however.

The species diversity of the entire stand was captured better by the
models than the diversity of tree regeneration, and only a few models
overpredicted stand-level diversity. Defining regeneration composition
as being proportional to adult abundance, regardless of productivity,
might be a simple and conservative assumption to safely maintain
relative species abundances [@hanbury-brown2022], although this may be
simplistic. Furthermore, based on the simulation results, there is no
evidence that models with feedback from the canopy captured better the
species diversity of the regeneration than those without feedback, with
the exception of ForCEEPS that significantly improved regeneration
species diversity with the model version that included feedback
(ForCEEPS(f)). The similar performance of models with and without
feedback is likely because the models put more weight on the
regeneration niche arising from abiotic and biotic filters, than from
the habitat niche of the adult trees [@grubb1977].

## Regeneration mortality

There are several factors that lead to mortality during the regeneration
phase, such as competition [@casper1997] and multiple abiotic factors
[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
expect that models that overestimate regeneration may have a
particularly high mortality between the 7 and 10 cm thresholds. However,
this was not consistently evident from the simulations. This implies
that these models must have an excess mortality in later stages, if we
assume that all models were able to capture the structure and
composition of the adult stands along the EuFoRIa gradient; this however
was not tested here.

Higher mortality towards the adult phase has important implications for
forest dynamics and the goods and services provided by forests. On the
one hand, mortality in later stages may erroneously enhance the share of
less shade tolerant species in the models [@klopcic2015], therefore
shifting the species composition. Unrealistic high stem densities for a
longer period of time may overestimate the role of tree regrowth in
carbon sink dynamics [@pugh2019]. Given our set of simulations and
analyses, we cannot conclusively assess what is happening in the models,
and further studies are required that focus on a wider range of tree
sizes and the fate of tree regeneration along such a size continuum.

## Model performance and model traits

In spite of the critical considerations above, it is remarkable that
most models did not deviate exceedingly from the observations with
respect to simulated regeneration levels -- after all, few if any of the
models' regeneration routines are sufficiently constrained by data, with
the exception of the empirical models xComp and SIBYLA. It is noteworthy
that in spite of their empirical basis, these two models did not match
empirical regeneration levels, in a similar magnitude as the other
models and even in opposite directions (SIBYLA: underestimation; xComp:
overestimation). It appears that using empirical data of limited
geographical scope and constrained to managed forests, as done in these
models, leads to extrapolation problems already under current climatic
conditions. For the other models, whose regeneration modules are not
strongly constrained by empirical data, the degrees of freedom that are
available for modeling regeneration [@konig_tree_2022] could have
implied that model performance would be much worse than what we found.

Our study showed that increasing complexity in the regeneration modules
is not linked with a higher accuracy of the projections of regeneration
levels, species composition or mortality at early tree stages, as there
was no significant relationship between model performance and model
complexity. Increasing complexity in regeneration modules has been
motivated by better process understanding and enhancing of model
accuracy [@bugmann2022]. However, more complex models do not necessarily
lead to better projections and rather to a reduced transparency and
lower predictive power [@franklin2020]. Thus, the question of what level
of detail is appropriate and parsimonious for modeling tree regeneration
processes remains open [@konig_tree_2022; @bugmann2022].

## Regeneration gradients on regeneration niches

Competition for light as a strong filter for tree regeneration has been
widely documented [@messier1999; @collet2006; @berdanier2016], but the
models examined here did not reproduce this expectation. However, it is
difficult to measure light availability at large spatial and temporal
scales. We used total stand basal area as a proxy for light availability
[cf. @schmid2021]. However, we were unable to consider light
availability restrictions caused by ground vegetation, which may be an
important filter for some EuForIa stands [@woltjer2008]. We found
pronounced differences in the stand basal area ranges simulated by the
models, but also between models and observations. This made it
impossible to evaluate the regeneration for the extremes of the stand
density ranges in some models. For example, regeneration levels at low
stand densities are relevant to assess how well forests are recovering
e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
the level of the simulated one-hectare samples, average basal area is
typically not very low as long as no larger disturbances occur, which
was explicitly excluded in our simulation protocol. Towards the other
end of the spectrum, i.e., with increasing stand basal area, it would be
reasonable to expect that the regeneration of the different species
would become sparser and drop out entirely under low-light conditions
[@klopcic2012; @zell2019]. However, few models showed this trend, thus
indicating that the relation between regeneration and light availability
is not yet captured correctly in most models. Yet, several of the models
that did feature an increase of regeneration with increasing basal area
include a feedback between seed production of mature trees and
regeneration. Thus, it seems that higher light competition does not
sufficiently compensate for increased seed availability with higher
basal area due to a higher abundance of mature trees.

There were pronounced differences how the main tree species were
represented by the models along the environmental gradients, in
particular the dominant species *Abies alba*, *Fagus sylvatica* and
*Picea abies*. The regeneration levels were sampled from simulations in
the equilibrium, and in this state it is expected that non-climax
species such as *Pinus sylvestris* or *Quercus* would be of minor
importance, or absent [@klopcic2015]. Most models captured this low
abundance, which is also found in the empirical data. Thus, although the
broad patterns are matched by many models, improvements in the
quantification of the regeneration niche of the species are needed, but
this cannot be done in the absence of robust datasets across multiple
stand development stages.

## Methodological considerations

The EuFoRIa data as used here are unique and well suited to better
understand tree regeneration. However, three aspects of these data may
represent considerable limitations. First, we made a comparison of tree
regeneration in an equilibrium state, but we cannot assess how close the
forests included in the EuFoRIa dataset were to such a state. The data
were collected in forest reserves where no management has taken place
for long periods of time. This makes our assumption of an equilibrium
between forest properties and environmental drivers more reasonable than
it might appear at first sight. In an analysis of primeval forests at
demographic equilibrium, @brzeziecki2021 found higher regeneration rates
than the ones observed in the EuFoRia data. Thus, the overestimation of
regeneration rates by the models may not be so problematic. Second, the
data were collected for regeneration above a 7 cm threshold, thus
limiting the assessment of tree regeneration to a holistic viewpoint.
This constituted a hard limit based on which we can understand only some
aspects of tree regeneration, which in its entirety often comprises a
rather long period since seed production [@price2001]. In reality, many
environmental constraints are acting on young trees [@käber2021] that we
were unable to assess. Yet, regeneration data with lower thresholds are
simply not available in a harmonized manner across large environmental
gradients. Third, the empirical data were collected from rather small
plots, while we sampled simulated regeneration levels from 1 ha areas,
which may lead to an incorrect representation of space. Even though the
strategy we adopted is not ideal, it represents a common challenge when
harmonizing diverse data sources originating from varied sampling
strategies [@portier2022]. It would have been extremely challenging for
such a variety of models to follow a protocol where the spatial sampling
size was different at each of the 200 sites, and it would have
introduced additional uncertainty in the results. Ultimately, we need to
acknowledge that the observed data represent just a snapshot of an
extremely complex and stochastic reality. This may at least partly
explain e.g. the lack of clear patterns in the simulation results along
the gradients of climate water balance and degree-days.

The design of our sampling protocol did not include spatial aspects such
as seed dispersal or detailed soil data. While we considered a wide
range of models of forest dynamics, from stand to global scales, the
simulation setup was limited to 1 ha regeneration values. This lack of
consideration of spatial scale is appropriate for stand-scale models,
but it potentially puts the landscape models at a disadvantage, as they
have been built to be accurate at the landscape level. Without spatial
context, we are limiting regeneration to the seed influx from the stand
itself, unless the model has a background seed input. Yet, the global
models should not be at a disadvantage due to the limited spatial
considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
-@lehsten2019], as they are usually lacking dispersal between cells and
are based on a strong abstraction of horizontal space
[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
available from the observed data, and independent, admittedly rather
coarse data for soil and climatic water balance had to be used instead.
It is noteworthy that many models represent drought based on detailed
indicators based e.g. on soil water holding capacity, which had to be
derived from a rough soil quality measure. This may at least partially
explain the unsatisfactory performance of many models along the drought
axis (i.e., climatic water balance).

## Research recommendations

With our study, we have demonstrated that models of forest dynamics need
a focus on tree regeneration modules to make them more robust. It
remains uncertain what level of detail is required to model regeneration
processes, and this must be addressed in future research. We recommend
that the improvement of the regeneration modules is implemented as
additional features that can be traced back, as done here for the
variants of ForClim and ForCEEPS, and that model complexity and
structure must always be connected with modeling objectives
[@albrich2020]. If it should be necessary to include more detail in the
regeneration models, this will come with higher parameterisation
efforts. This will most likely lead to lower generalization because the
required data will have to be collected from specific locations, as
currently there is no general, comprehensive regeneration dataset
available.

Therefore, we further recommend that more effort should be invested into
collecting harmonized datasets on tree regeneration. We emphasize that
datasets such as EuFoRIa are invaluable and should be expanded in both
their spatial extent (e.g., towards boreal and Mediterranean conditions)
as well as in time (e.g., continuing the monitoring into the future).
Such data will allow for a better evaluation of forest models and help
to reduce the uncertainty in their projections, which is crucial when
they are used as tools for predicting e.g. the impacts of anthropogenic
climate change.

In the present study, we have considered tree regeneration in the
equilibrium state only. It is equally important to understand how these
models project tree regeneration after changes in forest structure by
disturbances [@seidl2022], or under different management strategies
[@lindner2000]. However, this will require an entirely different set of
observed data, and potentially not all models of forest dynamics would
be able to assess the relationship of these aspects on tree
regeneration, e.g. due to the lack of disturbance or appropriate
management modules.

We recommend to investigate in detail the implications of the current
modeling strategies for tree regeneration and, ultimately, simulated
forest stand structure. This applies particularly to the erroneous
patterns of excess tree regeneration and later excess mortality, by
focusing on a wider range of tree sizes and the related regeneration
dynamics. We also recommend, especially for the landscape-level models,
the inclusion of explicit spatial considerations regarding tree
regeneration [@beckage2003]; this, however, is a serious challenge
regarding appropriate datasets, as inventory data have wide coverage,
but by definition do not allow for the assessment of spatial interaction
effects.

Exercises like the one presented here, where the models are operated in
"blind flight" mode, i.e., without the possibility of being tuned
towards capturing the expected patterns, should be repeated. Such
benchmarking exercises should next focus on aspects such as specific
model traits and the ecological formulations of particular
(sub-)processes, to better understand the implications of the
assumptions on which the models are based. Furthermore, the inclusion of
a wide range of models with different scales, type, population
structures, tree regeneration modules and complexity will ensure a wider
benefit to the whole modeling community.

## Conclusions

Models of forest dynamics are important tools in science and decision
support, and the formulation of tree regeneration has strong
implications for simulated forest properties. The 15 models and variants
used here are facing similar challenges in their representation of tree
regeneration: they generally overestimate tree regeneration levels, and
the simulated regeneration niche is not always captured accurately as a
function of biotic (light) and abiotic (temperature and moisture)
factors.

However, most models properly capture the diversity of the initial tree
community, and differences between model formulations, e.g., the
presence or absence of feedback from the adult trees, did not have a
strong effect for capturing the species composition of regeneration.

Regarding mortality in the early phase of tree life, many models that
feature a particularly high overestimation of regeneration levels are
compensating for this by a larger tree mortality. Often, this
compensation is not sufficient to reduce the high regeneration levels to
realistic values. Overall, there is no clear mortality pattern across
all models.

When capturing tree regeneration, the specific design decisions taken in
the development of any model are more important for its behavior
(accuracy) than scale (stand, landscape, global), modelling approach
(empirical vs. process-based), and complexity. Having both empirical and
process-based models in our set, the empirically-based models could have
been expected to have a better performance, as they were calibrated with
inventory data, but this was not the case. Similarly, higher model
complexity does not represent an improvement for capturing tree
regeneration.

Even though the regeneration routines of most of the models investigated
here have never been sufficiently constrained by robust data, the
regeneration projections are not overly off. This indicates that a lot
can be gained by a focus on the modeling of regeneration processes. The
representation of forest dynamics in these models would become much more
robust particularly in the face of climate change and post-disturbance
dynamics, thus strongly reducing the uncertainty in long-term
projections of future forest dynamics.

# Reproducibility and data availability statement

This study complies with the basic rules of reproducibility and has made
available the protocol of the study, the data and all the software that
has led to the findings presented in this paper. The protocol is
available at @díaz-yáñez2022. The code and data that support the
findings of this study are openly available at zenodo (*link to be
added*) and github (*link to be added*) [for the review process we have
provided these materials [on this
link](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password:
DownDataCodeRegen1)].

# Acknowledgements

This paper arises from a workshop held in Davos (Switzerland) in June
2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
Funding for the workshop was provided by COST Action CA19139 PROCLIAS
(PROcess-based models for CLimate Impact Attribution across Sectors),
supported by COST (European Cooperation in Science and Technology;
[https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research
Grant ETH-35 18-1). ODY was funded by the Swiss National Science
Foundation project 200020_188882 / 1. KB received support from the
European Research Council under the European Union's Horizon 2020
research and innovation program (Grant Agreement 101001905). JB
acknowledges funding from the long-term research development project no.
RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
financial support from I-Maestro (Innovative forest management
strategies for a resilient bioeconomy under climate change and
disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
project REINFORCE (Grant No. 891671). KM received funding from the
Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
no. ITMS2014+313011W580 supported by the Integrated Infrastructure
Operational Programme funded by the ERDF. CH acknowledges financial
support from Waldfonds WASIM2100 project. We would also like to thank
all the people that helped to make this study possible with
contributions at different stages such as Martin Gutsch and Giorgio
Vacchiano.

# References

```{=html}
```