### 2023-10-24 manuscript/manuscriptRW.Rmd
add revisions from Harald

---
subtitle: "\\fontsize{16pt}{3pt}\\selectfont \\textbf{\\textit{Tree regeneration in models of forest dynamics: a key priority for further research}} "
title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}"
author: 
- Olalla Díaz-Yáñez$^{*1}$, 
- Yannek Käber$^1$, 
- Tim Anders$^2$,
- Friedrich Bohn$^3$,
- Kristin H. Braziunas$^4$,
- Josef	Brůna$^5$,
- Rico Fischer$^3$,
- Samuel M. Fischer$^3$,
- Jessica Hetzer$^2$,
- Thomas Hickler$^2$,
- Christian	Hochauer$^6$,
- Manfred J. Lexer$^6$,
- Heike	Lischke$^7$,
- Paola	Mairota$^8$,
- Ján Merganič$^{9}$,
- Katarina Merganičová$^{10,11}$,
- Tobias Mette$^{12}$,
- Marco Mina$^{13}$,
- Xavier Morin$^{14}$
- Mats Nieberg,$^{15}$,
- Werner Rammer$^4$,
- Christopher P.O. Reyer$^8$,
- Simon	Scheiter$^2$,
- Daniel Scherrer$^{16}$,
- Harald Bugmann$^1$
date: $^1$ETH Zurich. Department of Environmental Systems Science. Institute of Terrestrial Ecosystems. Universitätstrasse 16, 8092 Zürich, Switzerland. \newline
      $^2$Senckenberg Biodiversity and Climate Research Centre (SBiK-F), 60325 Frankfurt am Main, Germany. \newline
      $^3$Department of Ecological Modelling, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
      $^4$Ecosystem Dynamics and Forest Management Group, TUM School of Life Sciences, Technical University of Munich, 85354 Freising, Germany. \newline
       $^5$Department of Geoecology, Institute of Botany of the Czech Academy of Sciences, Průhonice, Czech Republic. \newline
       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria \newline
       $^7$Dynamic Macroecology, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland. \newline
       $^8$Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, 14473 Potsdam, Germany. \newline
       $^9$Department of Soil Sciences, of Plants and Food, University of Bari Aldo Moro, Italy. \newline
       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen, Slovakia \newline
       $^{11}$Czech University of Life Sciences Prague, Faculty of Forestry and Wood Sciences, Kamýcká 129, 16500 Praha 6 – Suchdol, Czech Republic. \newline
       $^{12}$Department of Biodiversity of Ecosystems and Landscape, Slovak Academy of Sciences, Štefánikova 3, P.O.Box 25, 81499 Bratislava, Slovakia. \newline 
       $^{13}$Dpt. of Soil and Climate, Bavarian State Institute of Forestry, Germany. \newline 
       $^{14}$Institute for Alpine Environment, Eurac Research, Bolzano/Bozen, Italy. \newline 
       $^{15}$CEFE, CNRS Campus du CNRS 1919, route de Mende 34293 Montpellier cedex, France. \newline 
       $^{16}$Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland.  \newline
       $^*$Corresponding author (olalla.diaz@usys.ethz.ch)
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: no
    toc: no
    extra_dependencies: "subfig"
    latex_engine: xelatex
  word_document:
    reference_docx: template_paper.docx
fontsize: 12pt
linestretch: 2
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2007, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016, @langan2017, @scheiter2013 
linkcolor: blue
csl: chicago-author-date.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{font=scriptsize}
- \captionsetup[table]{font=scriptsize}
- \usepackage{titling}
- \predate{ }
- \postdate{  }
- \usepackage{fontspec}
- \setmainfont{Times New Roman}
---

```{r, echo=FALSE}
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries = FALSE)
  library(bibtex)
  options(tinytex.verbose = TRUE)
```

**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the
data and code are provided in an external repository to be evaluated
during the peer review process. The repository can be downloaded from
[this figshare private link:
https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
The protocol is available at @díaz-yáñez2022 ([with link:
https://osf.io/czdxp/](https://osf.io/czdxp/)). We will make publicly
available the code and data if the manuscript is accepted.

**Keywords:** tree regeneration \| tree recruitment \| tree
establishment \| forest dynamics \| models \| landscape \| stand \|
global

\newpage

# Abstract

Tree regeneration is a key process in forest dynamics, particularly in
the context of forest resilience and climate change. Models are pivotal
for assessing long-term forest dynamics, and they have been in use for
more than 50 years. However, there is a need to evaluate their capacity
to accurately represent tree regeneration. We assess how well current
models capture the overall abundance, species composition, and mortality
of tree regeneration.

Using 15 models built to capture long-term forest dynamics at the stand,
landscape, and global levels, we simulate tree regeneration at 200 sites
representing large environmental gradients across Central Europe. The
results are evaluated against extensive data from unmanaged forests.

Most of the models overestimate recruitment levels, which is compensated
only in some models by high simulated mortality rates in the early
stages of individual tree dynamics. Simulated species diversity of
recruitment generally matches observed ranges. Models simulating higher
stand-level species diversity do not feature higher species diversity in
the recruitment layer. The effect of light availability on recruitment
levels is captured better than the effects of temperature and soil
moisture, but patterns are not consistent across models. Increasing
complexity in the tree regeneration modules is not related to higher
accuracy of simulated tree recruitment. Furthermore, individual model
design is more important than scale (stand, landscape, global) and
approach (empirical, process-based) for accurately capturing tree
regeneration.

Despite the mismatches between simulation results and data, it is
remarkable that most models capture the essential features of the highly
complex process of tree regeneration, while not having been
parameterized with such data. We conclude that much can be gained by
evaluating and refining the modeling of tree regeneration processes.
This has the potential to render long-term projections of forest
dynamics under changing environmental conditions much more robust.

\newpage

# Introduction

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbance impacts and changing societal
demands on forest ecosystem services, it is becoming ever more important
to understand how forest structure, composition and function will
change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics have been developed over the past decades considering
the impacts of climate [@bugmann2022]. From these studies, it is evident
that we have a reasonably good understanding how to model tree growth
[@vanclay1997; @bugmann1996], and substantial efforts have been
dedicated to improving the representation of tree mortality
[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is much
less studied and is often represented rather coarsely in models
[@walck2011; @leishman1992; @price2001]. This presents an important
research gap, particularly in the context of climate-induced forest
disturbances and forest resilience.

Tree regeneration arises from multiple processes including pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree regeneration is
challenging. Currently, tree regeneration processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring them [as done in classical
forest growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2)
the use of a few simple environmental filters, as done in most forest
"gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994].

Overall, models are needed to (1) synthesize existing empirical data and
explore their relationships, (2) assess future tree regeneration e.g. in
the context of global change scenarios, and (3) identify the most
important processes that are shaping vegetation patterns. Given the
strategies that are used in models of forest dynamics to represent tree
regeneration, their behavior often is prone to problems, such as very
high levels of tree regeneration that necessitate excess mortality at
early stages of tree life to simulate correct stand structure and
composition [@kroiss2015]. Also, inadequately high species diversity in
tree regeneration may be simulated, which is characteristic of
"classical" forest gap models [@gutiérrez2016], at least as long as the
simulation set-up comprises a multitude of species. Some models use
calibration against local canopy-level data to constrain simulated tree
regeneration, which is likely to hamper the general applicability of
these models e.g. under scenarios of climate change. Furthermore,
correctly capturing the species composition of tree regeneration as a
function of the presence of seed trees in the canopy is often a
particular challenge, potentially leading to unrealistic successional
drift in the model, which must be corrected by factors that are hard to
parameterize [e.g., @lischke2006]. A related issue is the excessive
reduction of species diversity due to positive feedback effects, such
that eventually just single-species stands remain [@meier2011]. This is
sometimes corrected by the incorporation of a low level of seed influx
of all species at all times [@schumacher2006] or by restricting the
number of seeds per species in the seedbank [@lischke2006]. However,
simulated species composition is usually exceedingly sensitive to
assumptions about seed availability, whereas the parameters of such
functions are poorly constrained by field data. Lastly, there is often a
problem with insufficient observational constraints on parameter values
for models that start from very small tree sizes (e.g., 10 cm in height,
or even from seed) and track tree development in a process-oriented
manner by considering a multitude of ecological influences, rather than
emphasizing tree recruitment into a larger size class [e.g.,
@wehrli2006].

Modeling tree regeneration processes is challenging, and even in
empirical ecology, it has not received much attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail [@miina2006]. Yet, little data are available that cover all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Data on forest regeneration are often
fragmented, which constitutes a major problem for model building
[@clark1999a]. For example, monitoring on permanent plots (such as
National Forest Inventories) often measures the ingrowth of new trees
into a specific size class [@zell2019], however with a design that
captures the rate after the stand initiation phase [@hallsby2015]. Data
from permanent plots, although available across regions, are also highly
heterogeneous, rendering their use difficult in a modeling context
[@käber2021]. Lastly, targeted studies to measure tree regeneration on
experimental sites are often limited in spatial extent, species studied,
or the sub-set of processes that are investigated [@berdanier2016;
@collet2006].

Thus, a focus on the modeling of tree regeneration processes is sorely
needed and overdue [@price2001; @walck2011] if we are to make reliable
projections of future forest dynamics, i.e., when the models need to be
operated in extrapolation mode, as well as from a fundamental ecological
point of view for increasing systems understanding. In the present
study, for the first time, a large number of forest models commonly used
to assess forest dynamics under climate change are evaluated against a
continental-scale, multi-species harmonized dataset on tree recruitment
[@käber2023]. By tree recruitment, we refer to the passing of trees
across a specific diameter threshold ("ingrowth"). We included models
that are based on a range of "philosophies", from models operating at
the stand to the global scale as well as the range of models from
empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving and constraining this process --- including some that are not
incorporated explicitly in most models, such as deer browsing --- we do
not aim for a detailed statistical evaluation of each model. Instead, we
aim to evaluate the general recruitment patterns and magnitudes
simulated by the models, and benchmark the simulated regeneration niche
of multiple species against empirical data along a wide environmental
gradient of temperature, moisture, and light availability [@grubb1977].

More specifically, we aim to answer the following questions: (1) Are
models of forest dynamics capturing accurately tree recruitment levels,
initial tree species diversity and mortality in the recruitment? (2) Do
model traits explain differences in model performance? (3) How well do
the models capture total recruitment and the regeneration niches of
individual species across environmental gradients of light availability,
temperature, and soil moisture?

We evaluate the performance of the models by analyzing simulated data
alongside observed data. Specifically, we examine tree species
diversity, mortality rates in tree recruitment, and recruitment levels.
Then, we link model traits (e.g., complexity, scale) to performance, and
we assess whether modeled total recruitment or individual species
regeneration niches (i.e., proportional recruitment along an
environmental gradient) align with observational data.

# Material and methods

## Models

Fifteen models of forest dynamics (including two models featuring model
variants) developed for the stand, landscape or global scale were used
to simulate forest dynamics (Table \@ref(tab:tableModels)). The
approaches used for model construction and their origin differ strongly,
with most of the models featuring a largely "process-based" approach,
whereas two models are based on formulations derived from the
statistical analysis of inventory data (SIBYLA and xComp). Some of the
models largely rely on the original approach underlying forest gap
models (e.g., ForClim 1), whereas some are based on plant
ecophysiological processes (e.g., FORMIND and iLand). The different
approaches underlying the models have strongly influenced the
formulation of tree regeneration processes.

The models can be differentiated into 'regeneration' and 'recruitment'
approaches [@vanclay1994; @konig_tree_2022]. *Regeneration models*
include processes such as flowering and pollination, seed production,
seed dispersal, germination, and seedling growth, which ultimately lead
to the simulated number of established trees. *Recruitment models*, in
contrast, introduce a number of new trees with certain characteristics
such as biomass or diameter, without explicitly considering earlier
development processes. We can further distinguish models that feature a
feedback from stand properties to simulated recruitment, i.e., where the
level and species composition of recruitment is influenced by the adult
tree community via the production of seeds (or seedlings/saplings), from
models that do not contain such feedback.

The starting point for tree regeneration in the models differs as well,
ranging from seedbank, seed or seedling to sapling (i.e., trees much
larger than 10 cm height, often ca. 2-3 m tall). Models that start from
seed need to include a larger number of ecological processes such as
germination and survival of young seedlings, whereas models that start
from saplings have to aggregate via parameterizations several ecological
processes that are not treated explicitly. This latter approach reduces
model complexity but comes at the cost of blurred process
representation.

In the models used here, the overall complexity in the regeneration
modules varies considerably. Following @bugmann2022, we can classify
seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
TreeMig) as having rather high complexity in their regeneration modules
(mean regeneration formulation complexity across all processes \>0,
Table \@ref(tab:tableModels)), whereas the other eight models feture
relatively low complexity. Two models, ForClim and ForCEEPS, were used
here with two alternative variants of regeneration. ForClim variant 1
[@bugmann1996] is based on a regeneration module that adheres closely to
the concept introduced by @botkin1972 which is based on environmental
filters considering species-specific thresholds of light availability
and climatic variables. In contrast, ForClim variant 11 is adopting a
slightly more complex approach where individual species properties and
their relation to the environment are represented gradually along with
the relative suitability of different species (i.e., competition among
tree species) [@huber2020]. These two model variants allow us to
evaluate a more process-based and complex module (variant 11) against a
simple module (variant 1), while the rest of the model structure is
identical. Similarly, the two ForCEEPS variants allow us to isolate the
importance of the canopy feedback (i.e., simulated actual composition
and relative abundance of species in the plot) via seed trees for the
quantity and quality (e.g., diversity and composition) of simulated
regeneration, as one variant includes this feedback whereas the other
does not.

## Observed data

Recruitment data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel network of sites in forest reserves that represent the range of
environmental gradients in temperature and precipitation in Central
Europe as compiled in the EuFoRIa network [@euforia2019; @käber2023]
(Figure 1). The environmental coverage and the variety of forests
represented by this data is unprecedented within Europe. These forests
have been unmanaged for at least 10 years prior to their designation as
reserves, and most of them provide time series of natural forest
dynamics over multiple decades and up to ca. 85 years. The census
periods range from 3 to 37 years, with an average of 14 years. The data
provide information at the tree level, thus allowing for the sequential
comparison of processes such as individual-tree recruitment and death
between the measurements.

We selected 200 sites from this network as the benchmarking dataset for
the simulation, so as to be representative of the environmental
variation contained in the data. This was achieved by applying k-means
clustering to define 200 clusters of plots from the original set of 869
plots along the environmental dimensions of temperature, climatic water
balance, soil quality, slope and aspect. Each of these sites featured at
least two consecutive measurements. Recruitment size thresholds ranged
from stem diameters at breast height (dbh) from 0 to 10 cm. For our
study, we defined two datasets, where one included 165 sites with a
diameter threshold of 7 cm or lower, and the other included another 35
sites with diameter thresholds between 7 and 10 cm. Therefore we had in
total 200 sites with observations of newly recruited trees above 10 cm
and a subset of 165 sites with recruitment data above 7 cm. In the
results section, any observations from the data at the 7 cm threshold
originated from 165 sites.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2023 to provide recruitment
rates per ha and per decade. The observations used in this study
featured 30,900 newly established trees. Recruitment rates per site,
sample and decade ranged from zero to 1,246 trees, with a mean of 56
trees. Adult species composition was also available for each plot. For
more details on this dataset and the detailed recruitment information,
cf. @käber2023.

## Simulation protocol

The overarching goal of the simulation experiments was to assess tree
recruitment as it arises from empirical data against its representation
in a wide range of models of forest dynamics. We define tree recruitment
as the passing of a breast height diameter threshold of 7 and 10 cm
(synonym: ingrowth). To this end, each modeling group was provided with
a detailed protocol [@díaz-yáñez2022] with instructions how to perform
the simulations, input variables on climate and soil conditions, and the
list of expected output variables. Neither were further site information
(except for the data specified below) nor any data on tree recruitment
or forest stand features provided prior to the simulation.

The input variables were collected from different data sources and
aggregated to be suitable for the needs of the different models. Time
series of climatic variables from 1981 to 2018 were provided in hourly
[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
required for some models, such as relative humidity or vapor pressure
deficit, were calculated from these variables. The final instructions
for using the climate data were slightly different in each model, based
on the approach that best suited the model (Table
\@ref(tab:tableModels)). Soil quality data were provided as continuous
values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
The protocol also provided the elevation, slope and aspect for each of
the 200 sites [@asterscienceteam2019], but no other spatial information
such as coordinates, with the exception of iLand and aDGVM2, which
required blurred coordinates to derive highly detailed soil data.

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include a "background" disturbance
probability to increase chances of shade-intolerant species to
establish. The simulations were set up to sample species-specific
recruitment rates per decade and per ha in the equilibrium state of the
model, typically entailing a "spin-up" run (as we did not provide any
forest data). The modeling teams decided on the simulated area and how
they derived these samples. The exact length of the simulation was also
decided by the modeling teams (Table \@ref(tab:tableModels)). Further
details on how each modeling team prepared the simulations and the
outputs are available in Supplementary Material 1.

The simulations were run in the absence of management to a simulated
equilibrium ("Potential Natural Vegetation") with the current climate,
as the ultimate goal was to evaluate tree recruitment under comparable
and near-equilibrium conditions. This entails the assumption that (1)
the observations from the forest reserves reflect no traces of forest
management, and (2) there is an equilibrium between forest dynamics and
climate. While the former might be starting to be visible in many of the
EuFoRIA reserves, the latter may be more debatable. However, in the
absence of detailed data on the history of each plot in the EuFoRIa
network, some broad assumptions had to be made. Both the width of the
regeneration niche (i.e., in environmental space) as well as the
intensity of recruitment (i.e., the number of recruited trees per area
and per unit of time) were of interest.

The simulations were run for mixed-species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which regeneration data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L.,* and *Pinus
sylvestris L.* The same set of species was used at all 200 sites. Two
models included their standard set of species for the simulations, which
is much larger (i.e., ForClim 1, ForClim 11, TreeMig), and the species
simulated beyond the eleven species were lumped in a category "others".
In three models, fewer than these eleven species were simulated (4C,
xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). In 4C only *Fagus
sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp. and *Pinus
sylvestris* are parameterised. xComp simulations did not consider
*Fraxinus excelsior* due to a deprecated species parameterisation. In
LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not
parameterized and therefore these species could not be included in the
simulation. Finally, the model aDGVM2 does not simulate individual
species. Rather, community assembly processes and trait filtering
generate plant communities that are adjusted to the local biotic and
abiotic conditions, and the simulated plants can be classified into
ecological strategies based on their trait values in a post-processing
step.

Each of the models reported the number of trees crossing the two
diameter thresholds by sampling 200 times in a 10-year interval for each
species and per ha for each of the 200 sites. Multiple samples per site
were used to better understand the simulated variation within each site.
This was done using different strategies, depending on the model,
including (1) sampling simulated data from the same 1 ha plot in the
equilibrium over time, (2) sampling several 1 ha plots from the
simulated forest at one specific point in time (in the equilibrium) or
(3) a combination of (1) and (2). This resulted in 880,000 observations
per model that simulated the 11 species included in the protocol (200
sites, 200 samples per site, 11 species and two diameter thresholds).
For the models that simulated additional species, their regeneration
rates were aggregated as "others".

Two models did not provide results from all the simulated samples or
sites to avoid unrealistic results, as follows. In the model 4C, a
threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
unrealistic stand basal area data, and therefore not all the samples and
sites were reported. The reason for this is that 4C is not suitable for
long-term simulations without management, due to misrepresentations in
density-dependent mortality processes in long-term simulations and
assumptions of tree geometry that lead to unrealistic single tree
dimensions of very old individuals. LPJ-GUESS had 2% of the sample
outputs without tree recruitment, and these were considered as zero
stand basal area and zero regeneration for all the tree species
simulated; one site produced grassland rather than a forest and was not
included in the results.

## Data analysis

The simulation results were analyzed regarding (1) recruitment levels
(i.e., ingrowth number per unit time and space), (2) recruitment species
diversity, (3) recruitment mortality, (4) the relationship between model
performance and model traits, and (5) ingrowth gradients along the
regeneration niches. We evaluated species diversity across the models
and in relation to the observed data by calculating the Shannon index
$H$ based on the relative proportion of the species in terms of basal
area. It was calculated for the regeneration ($H_{R_{n}}$) (Equation
\ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
\ref{eqpiShannon}). The higher the value of the index, the higher the
species diversity at a particular site and sample.

```{=tex}
\begin{equation}
H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
\end{equation}
```
```{=tex}
\begin{equation}
p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
\end{equation}
```
```{=tex}
\begin{equation}
H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where $s$ is the total number of species present that have a basal area
larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
species $i$ in sample $n$ calculated as the regeneration basal area
($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
proportion of species $i$ calculated as the basal area of all trees
($BA_{n,i}$) of that species $i$ relative to total basal area
($Total_{BA_{n}}$) of the sample $n$. Species diversity was not assessed
for aDGVM2 as this model does not simulate individual species.

Mortality in tree recruitment was assessed based on the ratio of
recruitment between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
to estimate whether the ratio of recruitment between these two
thresholds was above or below the expected theoretical rate. The Reineke
self-thinning rule is usually calculated for even-aged, single species
stands and is based on a fixed relationship between the number of stems
and the quadratic mean diameter in fully stocked pure stands. The value
used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm
to be 77% higher than at 10 cm), calculated using Equation
\ref{eqReineke}.

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process-based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
between simulated and observed values of the recruitment levels and
species diversity by using paired t-tests.

Total recruitment and the regeneration niches of the individual species
were evaluated across the gradients of light availability, temperature,
and soil moisture as captured in the data. Stand basal area (the basal
area of all the trees in each sample above the recruitment threshold)
was used as a proxy for light availability at the forest floor, the
annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing
season warmth, and the climatic water balance as a proxy for soil
moisture [@speich2019]. Recruitment values of each species were
calculated as the mean across the 200 samples per site. The observed
data were modeled using a Generalized Additive Model [@wood2011] with a
negative binomial distribution and restricted maximum likelihood to
better understand the relationship between the environmental gradients
and the levels of observed regeneration, relative to the simulation
results per model.

In order to analyze the regeneration niches across the climatic
gradients, we focused on five common tree species or genera: *Abies
alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
*Quercus* spp. For these species, we calculated the share in the
recruited basal area per site, where each site represents a part of the
environmental space, as the mean across the available samples per site:

```{=tex}
\begin{equation}
\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
\end{equation}
```
where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
is the mean basal area in the recruitment (subsequently referred to as
"recruited basal area") of species $i$ at site $t$ across the available
samples at that site. Furthermore, regarding the share in the recruited
basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this
as zero when the recruited basal area of that species
($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited
basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited
basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.

# Results

## Recruitment levels

Simulated levels of recruitment varied strongly across the 15 models and
typically did not match the levels found in the forest reserve data
(Figure 2). Recruitment was overestimated in most models for both the 7
and 10 cm diameter thresholds, with the exception of the empirical stand
model SIBYLA, the landscape model Landis-II and the global model aDGVM2,
which estimated recruited levels at the lower end of the plausibility
interval of the observed data (Figure 2). The models with the largest
overestimation were the stand model PICUS and the landscape model
TreeMig. For most models, the variability of simulated recruitment
levels across the 200 sites (visible from the interquartile range in the
box plots of Figure 2) was similar to or smaller than observed, with the
notable exception of PICUS, where simulated recruitment variability was
much larger.

Both the observations and the simulated data had no recruitment in some
samples and at some sites (for details, cf. Table S1). The observed data
had 4% of the samples with no recruitment. Only three models had a
larger proportion of no recruitment (4C, Landis-II and aDVM2). Two
models always simulated recruitment for both the 7 and 10 cm threshold
(xComp and TreeMig), i.e., they did not feature any zero values. The
other ten models had a very low percentage of samples with no
recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences
of zeros than the observations.

## Tree species diversity of recruitment

Most models matched the level of diversity of the observed data quite
well (Figure 3). Five models overestimated recruitment diversity:
ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
particularly for the 7 cm diameter threshold. The model 4C is a special
case, as it simulated five species only, i.e., its diversity values are
not directly comparable to those of the other models, nor to the
observations. Only one model, Landis-II, consistently underestimated
recruited diversity.

In most models, there were only small but significant differences in the
species diversity of recruitment between 7 and 10 cm across sites
(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
maintained the recruitment diversity between the 7 and 10 cm thresholds
(i.e., the differences between them were not significant, see Table S2),
and the same was evident from the observed data.

For both the observed and the simulated data, and both recruitment
thresholds, species diversity in the recruitment compared to stand-level
diversity did not feature clear patterns across the models (Figures S1
and S2). Most models captured reasonably well or overpredicted species
diversity at the stand level (Figure 4A, left and center; cf. Figure
S3), but a characteristic feature was that the simulations had a much
lower variability of diversity compared to observations. Models
overpredicting stand-level species diversity included ForClim1, ForClim
11, FORMIND and TreeMig (Figure 4A, left). Several models underpredicted
species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A,
right).

The majority of the models overpredicted recruitment diversity; as
already observed in Figure 3, only two of them underestimated it (4C and
Landis-II) (Figure 4B and Figure S4). Both the models with and without
feedback from the adult trees to regeneration (via seed production)
overestimated, underestimated or captured reasonably well recruitment
species diversity compared to the observations. Again, simulated
diversity had considerably lower variability than observed diversity in
the recruitment.

## Recruitment mortality

There were strong differences among models regarding the mortality rate
between 7 and 10 cm (Figure 5). Most of the models featured a mortality
rate significantly larger than the observed data, thus at least
partially compensating for the general overestimation of recruitment
levels at 7 cm (Figure 2 and Table S3); six models did not feature a
significant difference compared to the observed data (ForCEEPS, FORMIND,
ForClim 1 and 11, iLand and aDGVM2). Four models (xComp, PICUS, LandClim
and TreeMig) featured very high mortality rates (i.e., well above the
Reineke self-thinning line), which compensated for the strong initial
overestimation of recruitment (cf. Figure 2). Interestingly, two models
that underestimated overall recruitment levels (SIBYLA and Landis-II;
cf. Figure 2) featured mortality rates that were close to but still
above the self-thinning line (Figure 5).

The fact that some models (and observed data) featured lower mortality
than expected by Reineke's self-thinning rule may point to facilitation,
or simply a higher mortality rate before the trees had reached 7 cm
diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
whose mortality was well above the self-thinning line, indicates that
they feature higher recruitment mortality compared to even-aged,
single-species forests. Some models yielded a mortality rate of
recruitment that is broadly compatible with the self-thinning rule
(e.g., 4C and LandClim).

The three models with the largest overestimation of the proportion of
recruitment at the 7 cm threshold featured the highest mortality between
the 7 and 10 cm thresholds (Figure 6A). The other models that
overestimated recruitment had mortalities lower than the expected
self-thinning ratio, with the exception of LandClim (Figure 6B). It is
noteworthy that most of the models that featured a low ratio of
recruitment between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11
and iLand) had only a small overestimation of recruitment at the 7 cm
threshold (Figure 5). aDGVM2 was the only model that underestimated
recruitment at the 7 cm threshold while featuring a low mortality
between the two thresholds (Figure 6C), and only a few models were close
to the correct recruitment levels at 7 cm while also being close to the
Reineke line (SIBYLA, Landis-II; cf. Figure 6D; and 4C as well as
LandClim).

Lastly, some models that had a low number of recruits at the 7 cm
threshold also had a small decrease of recruitment between 7 and 10 cm
(ratio 7-10 cm) (Figure S5). In these models, competition and
self-thinning are either not pronounced or must have occurred before the
trees had reached 7 cm. However, this pattern was not consistent across
models (Figure S5). It is noteworthy that there is a relationship
between the recruitment levels at 7 cm and the mortality rate (Table
S4): most of the models showed a positive effect (higher regeneration at
7 cm is coupled to an increase of the mortality between 7 and 10 cm).
While some models showed a negative effect, the linear trend was not
pronounced and not always significant regarding its slope.

## Model performance and model traits

There was no significant relationship between the mean complexity of the
regeneration module (Table \@ref(tab:tableModels)) and the
overestimation proportion at a diameter threshold of 7 cm (Figure S6).
All models except 4C had significant differences between the observed
and simulated mean recruitment values. Therefore it was not possible to
differentiate in terms of performance between models type (empirical or
process based), or their scales (stand, landscape or global) (Table S5).
The values of species diversity in the recruitment simulated by each
model were also assessed against the observed data to see if there was a
difference between models that include a feedback compared to those that
did not; all the models except ForCEEPs(f) and FORMIND had significant
differences (Table S6). It is noteworthy that for the model ForCEEPS the
pattern changed from the overprediction of species diversity in the
regeneration to a diversity level that is closer to observations when
the recruitment module included feedback (ForCEEPs(f)).

## Regeneration gradients and regeneration niches

When evaluating total tree recruitment levels along key gradients of
light availability (basal area), temperature (degree-days) and soil
moisture (climatic water balance), distinct features emerged: the models
reproduced the effect of basal area in both its magnitude and patterns
(Figure 7 and S7) considerably better than the effects of the climatic
gradients, where they featured varying patterns.

In the observed data, recruitment levels decreased clearly with
increasing stand basal area (Figure 7a). This trend was captured by the
Landis-II model only, albeit at much lower values of basal area than in
the observations. The other models featured distinctly different trends,
such as (1) an increase in recruitment levels with increasing total
basal area followed by the absence of recruitment at high values of
basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2)
almost constant recruitment levels with basal area (ForClim 11, xComp,
LandClim and aDGVM2), or (3) an increase of recruitment up to a certain
value of stand basal area followed by a decrease at even higher values
of stand basal area, with model-specific thresholds (4C, ForClim 1,
PICUS, SIBYLA and iLand).

The observed recruitment did not change much across the climatic water
balance gradient, and showed a slight increase with temperature. Neither
did the models generally match the observed recruitment patterns across
these gradients (Figure 7b,c), nor was there a clear pattern across
models. Regarding the soil moisture gradient, a group of models featured
decreasing recruitment with increasing soil moisture (4C, ForClim 1,
ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a
few models showed the opposite trend (ForCEEPS(f), FORMIND and
LPJ-GUESS). Across the temperature gradient, all the models that
featured a decrease in recruitment levels with increasing soil moisture
showed the opposite trend (i.e., more recruitment with increasing
temperature), except for 4C. The other models did not feature a clear
pattern.

Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
sylvestris*) and oak (*Quercus* spp.) were the species observed most
frequently in the EuFoRIa data. Most models captured well the share of
basal area in the recruitment of these five species compared to the
observed data along the two major climatic gradients (Figure 8, cf.
Figure S8). However, the models differed considerably in the way in
which they simulated these climatic niches.

First, some models overestimated recruitment across the entire gradients
of at least one of the five main tree species (Figure 8, e.g., ForCEEPS,
ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated
recruitment of more than one species for a large part of the climate
gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red colors in
Figure 8). These trends were consistent for both recruitment thresholds,
although the exact changes in the share of the recruited basal area were
different (Figure S8).

Second, there was some congruence in the simulation results by species
across the models. The share in the recruitment of Silver fir (*Abies
alba*) was captured well across the gradient by most models except for
xComp, PICUS, and LandClim, which overestimated the recruitment share of
this species. There were also some peculiarities evident for some
models, such as 4C that did not simulate *Abies alba*. The patterns
across models for *Fagus sylvatica* recruitment were more complex, as
its recruitment was underestimated by many models across the
environmental gradients featuring negligible regeneration at most sites,
while others overestimated it in different parts of the environmental
space. Most models overestimated the share of *Picea abies* recruitment
in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
spp. represented a small share of the recruitment in the observed data
mostly at warm-dry sites. Many models, however, erroneously featured
recruitment for these two species along most of the climatic gradients,
although with a low share of basal area. Five models (4C, FORMIND,
SIBYLA, xComp, and LandClim) had very small amounts of recruitment of
*Pinus sylvestris* and *Quercus* spp., or did not feature any
recruitment of these species at all.

Lastly, there is no model that performed well across the five species
regarding the species-specific recruitment levels in the environmental
space (Figure 8). Some models (e.g., ForClim11, PICUS) tended to
consistently overestimate the recruitment share of all five species, but
most models overestimated the recruitment share of some species while
underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).

# Discussion

Tree regeneration is a fundamental process in forest dynamics. Correctly
capturing it in dynamic models is fundamental to, for example, evaluate
post-disturbance dynamics and potential long-term recovery trajectories
as it will define the forest state in the first decades [@seidl2022]. If
this initial phase is not captured well, we can not properly assess
aspects such as forest resilience or the timing, magnitude and
progression of carbon sequestration. In this study, for the first time,
projections of tree recruitment from multiple models of forest dynamics
were confronted with a unique dataset from unmanaged forest reserves
across a large environmental gradient in Europe.

The EuFoRIa data [@käber2023] are exceptional, particularly with respect
to the number of records (number of sites and repeated measurements on
unmanaged forests), which is essential for capturing a highly 'noisy'
process such as tree regeneration. The use of this dataset for model
benchmarking provided novel insights on the ability of state-of-the-art
models to accurately simulate recruitment levels, its species
composition and mortality in an early stage of tree life, i.e., between
tree diameters of 7 and 10 cm. Overall, by adopting this approach a much
broader understanding resulted than if we had used these data for model
calibration: it is primarily from the shortcomings of the models that we
can gain ecological insights [cf. @trugman2022].

## Recruitment levels

Most of the models overestimated tree recruitment levels. This has
potentially far-reaching implications, e.g., regarding biomass (and thus
carbon) turnover, with a potential overestimation of the capacity for
forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
one specific stage, i.e., recruitment into the 7 and 10 cm diameter
classes. Trees at this size contribute little to carbon sequestration,
and if the excess regeneration at this stage is compensated -- in the
models -- soon thereafter by higher mortality, overall simulated tree
population dynamics may still be trustworthy at the level of stand
structure and attributes such as basal area or biomass.

From a modeling point of view, excess tree recruitment inevitably
requires excess mortality rates in a later stage, either -- as observed
for some models -- between the 7 and 10 cm diameter thresholds, or soon
after the 10 cm threshold has been crossed. In any case, correcting at
early stages for the expected forest densities at later stages is
equivalent to compensating for a first error (excess recruitment) by a
second error (excess mortality). It is highly likely that biased
projections will result, because the two errors are unlikely to be
perfectly linked and thus will not always compensate each other. Hence,
this structural problem of most models investigated here is problematic
particularly if the models are to be used under novel conditions such as
under climate change [e.g., @huber2021] or in a decision support context
[e.g., @thrippleton2021].

Recruitment levels define the structure and composition of future
forests, but it is equally important to correctly identify areas where
regeneration is lacking [@rammer2021]. There are multiple constraints to
the regeneration niche of tree species [@price2001], and therefore the
absence of regeneration is likely to be common [@fortin2007], even over
larger areas such as the one-hectare plots used here. Tree regeneration
data are characterized by zero-inflation. This was clearly evident from
the EuFoRIa dataset, but some of the models did not produce zeros at
all, or featured a very low proportion of zero data. This substantial
difference may be due to the fact that the simulation results were drawn
from equilibrium forests, whereas in reality many of the forest reserves
are recovering from past management activities and have become denser
over the past decades [e.g., @heiri2009], leading to less regeneration
than in an equilibrium situation. Another possible reason for this
difference may be the exclusion of factors like deer browsing or failure
to accurately measure establishment filters, such as competition with
the herb layer, site-specific or environmental limitations.

## Species composition of recruitment

Correctly capturing species composition in tree recruitment is important
to assess the future functional diversity of a forest, e.g., its
sensitivity to drought or resilience to disturbances [@redmond2015;
@seidl2022]. In the simulations, overall species diversity levels in the
recruitment were well within the observed range for half of the models
(7 out of 14). Thus, while most models are facing difficulties to
quantitatively match recruitment levels (cf. above), their performance
is better regarding the composition of recruitment as a function of
abiotic and biotic conditions. Most models maintained or even decreased
recruitment diversity between the 7 and 10 cm thresholds, and the same
is visible from the empirical data, but the differences between the 7
and 10 cm thresholds were not significant. Diversity variations may be
explained by the way the individual models consider regeneration
processes [@konig_tree_2022], e.g., regeneration can be linked to the
seed/seedling bank, seed rain from dispersal or from a pool of available
species.

The species diversity of the entire stand was captured better by the
models than the diversity of tree recruitment, and only a few models
overpredicted stand-level diversity. Defining recruitment composition as
being proportional to local adult abundance, regardless of productivity,
might be a simple and conservative assumption to maintain relative
species abundances [@hanbury-brown2022], although this may be
simplistic. Furthermore, based on the simulation results, there is no
evidence that models with feedback from the canopy (in terms of species
composition of regeneration) captured the species diversity of the
recruitment better than those without feedback, with the exception of
ForCEEPS that featured significantly more accurate recruitment species
diversity with the model version that included feedback (ForCEEPS(f)).
The similar performance of models with and without feedback is likely
because the models put more weight on the regeneration niche arising
from abiotic and biotic filters, than from the habitat niche of the
adult trees [@grubb1977].

## Recruitment mortality

There are several factors that lead to mortality during the regeneration
phase, such as competition [@casper1997] and multiple abiotic factors
[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
expect that models that overestimate recruitment at 7 cm may have a
particularly high mortality between the 7 and 10 cm thresholds. However,
this was not consistently evident from the simulations. This implies
that these models must have an excess mortality in later stages, if we
assume that all models were able to capture the structure and
composition of the adult stands along the EuFoRIa gradient; this however
was not tested here.

Higher mortality towards the adult phase has important implications for
forest dynamics and the goods and services provided by forests. On the
one hand, mortality in later stages may erroneously enhance the share of
less shade tolerant species in the models [@klopcic2015], therefore
shifting the species composition. Unrealistic high stem densities for a
longer period of time may overestimate the role of tree regrowth in
carbon sink dynamics [@pugh2019]. Given our set of simulations and
analyses, we cannot conclusively assess what is happening in the models,
and further studies are required that focus on a wider range of tree
sizes and the fate of tree regeneration along such a size continuum.

## Model performance and model traits

In spite of the critical considerations above, it is remarkable that
most models did not deviate exceedingly from the observations with
respect to simulated recruited levels -- after all, few if any of the
models' regeneration routines are well constrained by data, with the
exception of the empirical models xComp and SIBYLA. It is noteworthy
that in spite of their empirical basis, these two models did not match
empirical recruitment levels, in a similar magnitude as the other models
and even in opposite directions (SIBYLA: underestimation; xComp:
overestimation). It appears that using empirical data of limited
geographical scope that are constrained to managed forests, as done in
these models, leads to extrapolation problems already under current
climatic conditions. For the other models, whose regeneration modules
are not strongly constrained by empirical data, the multiple strategies
that are available for modeling regeneration processes
[@konig_tree_2022] could have implied that model performance would be
much worse than what we found.

Our study showed that increasing complexity in the regeneration modules
is not linked to a higher accuracy of the projections of recruitment
levels, species composition or mortality at early tree stages, as there
was no significant relationship between model performance and model
complexity. Increasing complexity in regeneration modules has been
called for a long time ago (e.g., @price2001) motivated by better
process understanding and with the goal of enhancing model accuracy
[@bugmann2022]. However, more complex models do may not necessarily lead
to better projections but rather to reduced transparency and lower
predictive power [@franklin2020]. Thus, the question regarding the level
of detail that is appropriate and parsimonious for modeling tree
regeneration remains open [@konig_tree_2022; @bugmann2022].

## Regeneration gradients on regeneration niches

Competition for light as a strong filter for tree regeneration has been
widely documented [@messier1999; @collet2006; @berdanier2016], but the
models examined here did not reproduce this expectation (i.e.,
decreasing recruitment levels with increasing stand basal area).
However, it is difficult to measure light availability at large spatial
and temporal scales. We used total stand basal area as a proxy for light
availability [cf. @schmid2021]. Yet, we were unable to consider light
availability restrictions caused by ground vegetation, which may be an
important filter for forest dynamics at least in some EuFoRIa stands
[@woltjer2008]. We found pronounced differences in the ranges of stand
basal area simulated by the models, but also between models and
observations. This made it impossible to evaluate tree recruitment for
the extremes of the stand density ranges in some models. This is
unfortunate because recruitment levels at low stand density are relevant
to assess how well forests are recovering e.g. after gap creation due to
disturbance [@seidl2022; @grubb1977]. At the level of the simulated
one-hectare samples, average basal area is typically not very low as
long as no larger disturbances occur, which was explicitly excluded in
our simulation protocol. Towards the other end of the spectrum, i.e.,
with increasing stand basal area, it would be reasonable to expect that
the recruitment of the different species would become sparser and drop
out entirely under low-light conditions [@klopcic2012; @zell2019].
However, few models showed this trend, thus indicating that the relation
between regeneration and light availability is not yet captured
correctly in most models. Yet, several of the models that did feature an
increase of regeneration with increasing basal area include a feedback
between seed production of mature trees and regeneration. Thus, it seems
that in these models higher light competition does not sufficiently
compensate for increased seed availability with higher basal area due to
a higher abundance of mature trees.

Lastly, there were pronounced differences how the main tree species were
represented by the models along the environmental gradients, in
particular the dominant species *Abies alba*, *Fagus sylvatica* and
*Picea abies*. The recruitment levels were sampled from simulations in
the equilibrium, and in this state it is expected that non-climax
species such as *Pinus sylvestris* or *Quercus* would be of minor
importance, or absent [@klopcic2015]. Most models captured this low
abundance, which is also found in the empirical data. Thus, although the
broad patterns are matched by many models, improvements in the
quantification of the regeneration niche of the species are needed, but
this cannot be done in the absence of robust datasets across multiple
species.

## Methodological considerations

The EuFoRIa data as used here are well suited to better understand tree
regeneration. However, three aspects of these data may represent
considerable limitations. First, we made a comparison of tree
regeneration in an equilibrium state, but we cannot assess how close the
forests included in the EuFoRIa dataset were to such a state. The data
were collected in forest reserves where no management has taken place
for long periods of time. This makes our assumption of an equilibrium
between forest properties and environmental drivers more reasonable than
it might appear at first sight. In an analysis of primeval forests at
demographic equilibrium, @brzeziecki2021 found higher recruitment rates
than the ones observed in the EuFoRia data. Thus, the overestimation of
recruitment rates by the models may not be so problematic. Second, the
data were collected for recruitment above a 7 cm threshold, thus
limiting the assessment of tree regeneration to a specific point of
stand dynamics. This constituted a hard limit based on which we can
understand only some aspects of tree regeneration, which in its entirety
often comprises a rather long period since seed production [@price2001].
In reality, many environmental constraints are acting on young trees
[@käber2021] that we were unable to assess. Yet, recruitment data with
lower thresholds are simply not available in a harmonized manner across
large environmental gradients. Third, the empirical data were collected
from rather small plots, while we sampled simulated recruitment levels
from 1 ha areas, which may lead to an incorrect representation of space.
Even though the strategy we adopted is not ideal, it represents a common
challenge when harmonizing diverse data sources originating from varied
sampling strategies [@portier2022]. It would have been extremely
challenging for such a variety of models to follow a protocol where the
spatial sampling size was different at each of the 200 sites, and it
would have introduced additional uncertainty in the results.

The design of our sampling protocol did not include spatial aspects such
as seed dispersal or detailed soil data. While we considered a wide
range of models of forest dynamics, from stand to global scales, the
simulation setup was limited a spatial scale of 1 ha. This lack of
consideration of spatial scale relationships is appropriate for
stand-scale models, but it potentially puts the landscape models at a
disadvantage, as they have been built to be accurate at the landscape
level: without any spatial context, we are limiting tree regeneration to
the seed influx from the stand itself, unless the model has a background
seed input. Yet, the global models should not be at a disadvantage here
due to their inherently limited spatial considerations [but cf. Snell et
al. -@snell2018; or Lehsten et al. -@lehsten2019], as they are usually
lacking dispersal between cells and are based on a strong abstraction of
horizontal space [@hanbury-brown2022]. Lastly, detailed data on soil
conditions were not available from the observed data, and independent,
admittedly coarse data for soil properties and the climatic water
balance had to be used instead. It is noteworthy that many models
represent drought using detailed indicators based e.g. on soil water
holding capacity, which had to be derived from a rough soil quality
measure. This may at least partially explain the unsatisfactory
performance of many models along the drought axis (i.e., climatic water
balance).

## Research recommendations

With our study, we have demonstrated that models of forest dynamics need
a focus on their regeneration modules to make them more robust. It
remains uncertain what level of detail is required to model tree
regeneration, and this must be addressed in future research. We
recommend that the improvement of the regeneration modules is
implemented as additional features that can be traced back, as done here
for the variants of ForClim and ForCEEPS, and that model complexity and
structure must always be connected with modeling objectives
[@albrich2020]. If it should be necessary to include more detail in the
regeneration models, this will come with higher parameterisation
efforts. This will most likely lead to lower generalization because the
required data will have to be collected from specific locations, as
currently there is no general, comprehensive regeneration dataset
available.

Therefore, we further recommend that more effort should be invested into
collecting harmonized datasets in a site-specific manner covering the
different aspects leading to tree regeneration. We emphasize that
datasets such as EuFoRIa are invaluable and should be expanded in both
their spatial extent (e.g., towards boreal and Mediterranean conditions)
as well as in time (e.g., continuing the monitoring into the future).
Such data will allow for a better evaluation of forest models and help
to reduce the uncertainty in their projections, which is crucial when
they are used as tools for predicting e.g. the impacts of anthropogenic
climate change.

In the present study, we have considered tree regeneration in the
equilibrium state only. It is equally important to understand how these
models project tree regeneration after changes in forest structure by
disturbances [@seidl2022], or under different management strategies
[@lindner2000]. However, this will require an entirely different set of
observed data, and potentially not all forest models would be able to
assess the relationship of these aspects on tree regeneration, e.g. due
to the lack of disturbance or appropriate management modules.

We recommend to investigate in detail the implications of the current
modeling strategies for tree regeneration and, ultimately, simulated
forest stand structure. This applies particularly to the erroneous
patterns of excess tree regeneration and later excess mortality, by
focusing on a wider range of tree sizes and the related regeneration
dynamics. We also recommend, especially for the landscape-level models,
the inclusion of explicit spatial considerations regarding tree
regeneration [@beckage2003]; this, however, is a serious challenge
regarding appropriate datasets, as large-scale inventory data have a
wide coverage, but by definition do not allow for the assessment of
spatial interactions.

Exercises like the one presented here, where the models are operated in
"blind flight" mode, i.e., without the possibility of being tuned
towards capturing the expected patterns, should be repeated. Such
benchmarking exercises should next focus on aspects such as specific
model traits and the ecological formulations of particular
(sub-)processes, to better understand the implications of the
assumptions on which the models are based. Furthermore, the inclusion of
a wide range of models with different (1) scales, (2) approaches for
capturing population structure, (3) tree regeneration modules and (4)
complexity of their formulations will ensure a large benefit to the
entire modeling community and beyond.

## Conclusions

Models of forest dynamics are important tools in science and decision
support, and the formulation of tree regeneration has strong
implications for simulated forest properties. The 15 models and variants
used here are facing similar challenges in their representation of tree
regeneration: they generally overestimate tree recruitment levels, and
the simulated regeneration niche is not always captured accurately as a
function of biotic (light) and abiotic (temperature and moisture)
factors.

However, most models properly capture the diversity of the initial tree
community, and differences between model formulations, e.g., the
presence or absence of feedback from the adult trees did not have a
strong effect on capturing the species composition of regeneration.

Regarding mortality in the early phase of tree life, many models that
feature a particularly high overestimation of recruitment levels are
compensating for this by a larger tree mortality. Often, this
compensation is not sufficient to reduce the high recruitment levels to
realistic values. Overall, there is no clear mortality pattern across
all models.

When capturing tree regeneration, the specific design decisions taken in
the development of any model are more important for its behavior
(accuracy) than scale (stand, landscape, global), modelling approach
(empirical vs. process-based), and complexity. Having both empirical and
process-based models in our set, the empirically-based models could have
been expected to have a better performance, as they were calibrated with
inventory data, but this was not the case. Similarly, higher model
complexity does not represent an improvement for capturing tree
regeneration.

Even though the regeneration routines of most of the models investigated
here have never been constrained well by robust data, their projections
of recruitment are not overly off. This indicates that a lot can be
gained by a focus on the modeling of regeneration processes. The
representation of forest dynamics in these models would become much more
robust particularly in the face of climate change and post-disturbance
dynamics, thus strongly reducing the uncertainty in long-term
projections of future forest dynamics.

# Author contributions

O. Díaz-Yáñez and Y. Käber contributed equally and share the first
authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
design. Y. Käber and O. Díaz-Yáñez prepared the input data used by all
modeling teams. All authors contributed by conducting the simulations
with their respective models. O. Díaz-Yáñez led the analysis of the
simulation results with contributions during the workshop from Y.Käber,
T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T.
Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X.
Morin, W. Rammer, and D. Scherrer. O. Díaz-Yáñez and H. Bugmann led the
writing of the manuscript and provided the first draft. All authors
participated in the revision of the manuscript and approved its
submission.

# Acknowledgements

This paper arises from a workshop held in Davos (Switzerland) in June
2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
Funding for the workshop was provided by COST Action CA19139 PROCLIAS
(PROcess-based models for CLimate Impact Attribution across Sectors),
supported by COST (European Cooperation in Science and Technology;
[https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research
Grant ETH-35 18-1). ODY was funded by the Swiss National Science
Foundation project 200020_188882 / 1. KB received support from the
European Research Council under the European Union's Horizon 2020
research and innovation program (Grant Agreement 101001905). JB
acknowledges funding from the long-term research development project no.
RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
financial support from I-Maestro (Innovative forest management
strategies for a resilient bioeconomy under climate change and
disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
project REINFORCE (Grant No. 891671). KM received funding from the
Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
no. ITMS2014+313011W580 supported by the Integrated Infrastructure
Operational Programme funded by the ERDF. CH acknowledges financial
support from Waldfonds WASIM2100 project. We would also like to thank
all the people who helped to make this study possible with contributions
at different stages particularly Martin Gutsch and Giorgio Vacchiano.

# Conflict of interest statement

The authors declare no conflicts of interest.

# References

```{=html}
```
::: {#refs}
:::

\newpage

# Table

```{r tableModels, echo=FALSE,  message=FALSE, warning=TRUE}
library(knitr)
modelDF <- readr::read_csv(here::here("manuscript", "modelsDescription.csv"))
modelDF$Sampling <- NULL

#footnotes
names(modelDF)[2] <- paste0(names(modelDF)[2], "(a)")
names(modelDF)[3] <- paste0(names(modelDF)[3], "(b)")
names(modelDF)[4] <- paste0(names(modelDF)[4], "(c)")
names(modelDF)[5] <- paste0(names(modelDF)[5], "(d)")
names(modelDF)[6] <- paste0(names(modelDF)[6], "(e)")
names(modelDF)[11] <- paste0(names(modelDF)[11],"(f)")

kable(modelDF,
      format = "latex", booktabs = TRUE,
      caption = "") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |>
      kableExtra::column_spec(column = 2, width = "3em") |>
      kableExtra::column_spec(column = 3, width = "4em") |>
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "6em") |>
      kableExtra::column_spec(column = 7, width = "5em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "8em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "10em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical",
                                 "C-Cohort, I-Individual",
                                 "Y-Yes, N-No",
                                 "R-Regeneration, Rc-Recruitment",
                                 "D-Daily, M-Monthly"),

                 notation = "alphabet")
```

\newpage

# Table and Figure captions

**Table 1.** Models included in this study, their characteristics and
simulation strategies used. *Scale* refers to the models scale, *Type*
to the models type, *Population structure* to the model population
approach, *Feedback* to the inclusion or not of adult tree feedback to
the level and species composition of regeneration, *Approach* to the
model regeneration module approach, *Start from* to the stage from which
trees are recruited, *Complexity*, to the mean regeneration formulation
as estimated in (Bugmann & Seidl, 2022), *Species* to which species from
the original eleven species were simulated, *Runtime* (spin up) [years]
to the number of years used per model in their simulations for the spin
up period and regeneration sampling; *Climate data type (use of years)*
to the type of data used in the simulations and how it was used, and
*Reference* provides the link to the publication with more model
details. Further information can be found on the individual model
reports available in the supplementary materials.

**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used
in the present study, the colors show the dominant tree species per site
in the observed regeneration at the 10 cm threshold.

**Figure 2.** Mean regeneration levels across all samples per site,
plotted for the 200 sites and for each model. The red dashed lines shows
the 25th and 75th percentiles for the 7 cm diameter threshold in the
observed data. There are two boxplots for each model where the left
lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
cm

**Figure 3.** Shannon index of tree regeneration, calculated by basal
area, as the mean value across the 200 samples per site for the diameter
thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
25th and 75th percentiles for the 7 cm diameter threshold in the
observed data.

**Figure 4.** Mean Shannon index across all samples per site for
observed and simulated data. Each plot shows one pattern represented by
one exemplary model of each category (overpredicted, intermediate and
underpredicted). The full data with the grouping of the models are shown
in Figures S3 and S4. n indicates the number of models falling in each
group. A) Examples for the three trends across models for species
diversity at the stand level. B) Examples for the three trends across
models for regeneration at the 7 cm threshold (165 sites).

**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm
thresholds. Dashed blue lines mark a ratio equal to 1, indicating no
decrease in tree regeneration between 7 and 10 cm, and a ratio equal to
1.77, corresponding to the Reineke self-thinning ratio under even-aged
conditions.

**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm
thresholds (regeneration 7 cm / regeneration 10 cm) and the
overestimation proportion at 7 cm ([simulated - observed] / observed)
for the mean regeneration per model across sites and samples. The
horizontal dashed lines show a ratio equal to 1, indicating no decrease
of regeneration between 7 and 10 cm, and a ratio equal to 1.77
corresponding to the Reineke self-thinning line under even-aged
conditions. The vertical line indicates zero overestimation.

**Figure 7.** Mean recruitment across the 200 samples per site with the
y-axis scaled differently by model, for the 200 sites against gradients
of (a) total basal area; (b) climatic water balance; (c) seasonal
degree-day sum. The values were split into ten bins; the red lines
represent a GAM model showing the trend in the observed data. Y-axis are
not equally scaled.

**Figure 8.** Regeneration levels of the five main species in terms of
their basal area share in the observations (top row) and the performance
of each model across the environmental gradients (other rows) for the 7
cm regeneration threshold. The values shown are the mean of the 200
samples per site and across the sites in each bin (tile), with ten bins
per gradient. The sizes of the circles represent the ratio between the
regeneration basal area of the species and the total regeneration basal
area for all species. The absence of a circle indicates a zero basal
area share in the regeneration, or the absence of regeneration
altogether. The color gradient (for the models only) shows the
difference between the simulated and observed ratio of regeneration
basal area of the species and the total regeneration basal area for all
the species.

\newpage