### 2023-10-31 manuscript/review/editorsLetter.qmd
editors latter notes

---
title: "letterEditor_revision"
format: pdf
editor: visual
---

Dear Dr. Charles Canham,

I would like to express my sincere gratitude for taking the time to review our manuscript. Your insightful feedback on our findings was immensely helpful.

We have taken all the reviewer\'s comments and suggestions into consideration and have made significant changes to the manuscript. Please find enclosed a detailed report outlining the actions we have taken in response to each of the reviewer's comments.

Thank you again for your time and effort. We look forward to hearing from you soon.

Best regards,

Olalla Díaz-Yáñez, on behalf of all the co-authors