### 2024-02-12 manuscript/review/editorsLetter.qmd
last comment not pushc from the editors revision

---
title: "letterEditor_revision"
format: pdf
editor: visual
---

Dear Dr. Charles Canham,

Thank you for taking the time to review our manuscript and for your valuable feedback.

In this revised version, we have taken all the reviewer's comments and suggestions into consideration and have made significant changes to the manuscript accordingly. We have submitted a detailed report outlining the actions we have taken in response to each of the reviewer's comments. We believe that the changes we have made have significantly improved the quality of our manuscript.

Thank you again for your time. We look forward to hearing from you soon.

Best regards,

Olalla Díaz-Yáñez, on behalf of all the co-authors