### 2024-02-12 manuscript/review/editorsLetter.qmd

---
title: "letterEditor_revision"
format: pdf
editor: visual
---

Dear Dr. Charles Canham,

Thank you for taking the time to review our manuscript and for your valuable feedback.

We have taken all the reviewer\'s comments and suggestions into consideration and have made significant changes to the manuscript. Please find enclosed a detailed report outlining the actions we have taken in response to each of the reviewer's comments.

Thank you again for your time. We look forward to hearing from you soon.

Best regards,

Olalla Díaz-Yáñez, on behalf of all the co-authors